---
title: "PC1Analysis"
author: "Qianhui"
date: "12/17/2018"
output: html_document
---

This document is mainly for part6 of the GEO data set analysis, that is pull out PC1 probes.


```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", error = TRUE, echo = TRUE) # fig.width = 7, 
```


# 1. load data and packages
```{r package, warning=FALSE, message=FALSE}
library(tidyverse)
library(plyranges)
library(readxl)
library(minfi)
library(magrittr)
library(tibble)
library(data.table)
library(limma)
library(ewastools)
library(ENmix)
library(wateRmelon)
library(WGCNA)
library(BSgenome.Hsapiens.UCSC.hg19)
library(missMethyl)
library(FactoMineR)
library(doParallel)
library(ggplot2)
library(gridExtra)
library(ggpubr)
library(RColorBrewer)
pal <- brewer.pal(8, "Dark2")
# GEOmeta, GEO_phenotypes
load(file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_phenotypes_add10.RData")

# Join.df_v2, GEO_phenotypes_joinedSamples, MDS_joinedSamples
load(file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_joinedSamples551_Beta_MDS.RData")

# # PCA_Placenta_M, PCA_PlacentaT1_M, PCA_PlacentaTerm_M
# load(file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/PCA_placenta_deciduals_allT1Term_BMval.RData")

PCA_Placenta_M <- readRDS(file  = "/home/qianhui/DNAme/Process_decidual/RDSfiles/PCA_all.Rds")

# annotation needed for using functions
ann450k <- readRDS(file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/ann450k_GEO15sets.rds")
annEPIC <- readRDS(file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/annEPIC_GEO15sets.rds")

# BM list after normalisation, but not corrected for batch across samples
# NormBMlist <- readRDS(file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/NormBMlist.rds")

NormBMlist <- readRDS(file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/NormBMlist_300k.rds")

CutOffRight <- quantile(PCA_Placenta_M$ind$coord[, 1], 0.75)+ 1.5*IQR(PCA_Placenta_M$ind$coord[, 1])
CutOffLeft <- quantile(PCA_Placenta_M$ind$coord[, 1], 0.25)- 1.5*IQR(PCA_Placenta_M$ind$coord[, 1])

CutOffRightNames <- PCA_Placenta_M$ind$coord[, 1][PCA_Placenta_M$ind$coord[, 1]>CutOffRight] %>% sort() %>% names
CutOffLeftNames <- PCA_Placenta_M$ind$coord[, 1][PCA_Placenta_M$ind$coord[, 1]<CutOffLeft] %>% sort()%>% names

# M values and matched phenodata
registerDoParallel(cores=5)

table(colnames(NormBMlist$NormMAll) %in% GEO_phenotypes$Sample_name)

Matched_pd <- GEO_phenotypes[match(colnames(NormBMlist$NormMAll), GEO_phenotypes$Sample_name),]
Matched_pd$Outlier <- ifelse(Matched_pd$Additional_Info%in%"Outlier"|Matched_pd$Sample_name%in%c("GSM1702177_7970368142_R05C02", "GSM1947213_6008581028_R01C01"), "Outlier", "Standard")


# PMDs
PMDs_hg19 <- readRDS(file = "/home/qianhui/DNAme/data/Placenta9-PMDpublished_hg19GR.unlisted.rds")

# ICRs
ICRsGR <- readRDS(file ="/home/qianhui/DNAme/Process_decidual/ICRs/ICRsGR.rds")

```


# the contribution of probes to PC1 (all samples including different tissue types)

## PCA plot: get p-values for probes contributing to PC1, and plot scree plot

```{r}
DimPval <- dimdesc(PCA_Placenta_M, axes = 1:2)
DimPval$Dim.1$quanti[,1] %>% quantile()
DimPval$Dim.1$quanti[,2] %>% quantile()

ProbeCorAndPval_PC1 <- DimPval$Dim.1$quanti %>% 
  as.data.frame() %>% 
  rownames_to_column(var="ProbeNames") %>% 
  `colnames<-`(c("ProbeNames", "CorrelationPC1", "p.valuePC1"))

### plot eigen values for each PCA
barplot(PCA_Placenta_M$eig[,1],main="Eigenvalues",names.arg=1:nrow(PCA_Placenta_M$eig))


# scree plot for PCAs
eigenvalues <- PCA_Placenta_M$eig[1:10, ]
barplot(eigenvalues[, 2], names.arg=1:nrow(eigenvalues), 
       main = "Variances",
       xlab = "Principal Components",
       ylab = "Percentage of variances",
       col ="steelblue")
# Add connected line segments to the plot
lines(x = 1:nrow(eigenvalues), eigenvalues[, 2], 
      type="b", pch=19, col = "red")

```


## get top 2% probes in PC1
```{r Contribution2PC1}

# contribution of probs in Dim1
ProbeContrib.df <- PCA_Placenta_M$var$contrib %>% as.data.frame() %>% 
  rownames_to_column(var="Name")
###### quick look at conribution percentages:
summary(ProbeContrib.df)
###### get Dim1 top 2% probes by ordering the contribution percentages of Dim1
ProbeContrib.df.orderedTop2perc <- ProbeContrib.df[base::order(-ProbeContrib.df$Dim.1),] %>% 
   dplyr::slice(1:(0.02*nrow(ProbeContrib.df))) %>% 
   dplyr::left_join(ProbeCorAndPval_PC1, c("Name"="ProbeNames"))

ProbeContrib.df.orderedTop2perc$CorrelationPC1 %>% quantile()

ProbeContrib.df.orderedTop2perc$p.valuePC1 %>% quantile()

```


Calculate beta difference between placenta and mixed placenta sampels:

```{r}

Matched_pd_placenta <- Matched_pd[Matched_pd$Sample%in%"Placenta",]

NormBAll_placenta <- NormBMlist$NormBAll[,colnames(NormBMlist$NormBAll)%in%Matched_pd_placenta$Sample_name]

Placenta_names <- Matched_pd_placenta[Matched_pd_placenta$Outlier%in%"Standard",]$Sample_name

Outlier_names <- Matched_pd_placenta[Matched_pd_placenta$Outlier%in%"Outlier",]$Sample_name


# for all probes, delta beta
Placenta_MeanB_allProbes <- NormBAll_placenta[,colnames(NormBAll_placenta)%in%Placenta_names] %>% 
  rowMeans()

Outlier_MeanB_allProbes <- NormBAll_placenta[,colnames(NormBAll_placenta)%in%Outlier_names] %>% 
  rowMeans()
  

table(abs(Outlier_MeanB_allProbes-Placenta_MeanB_allProbes) >= 0.2)

table(Outlier_MeanB_allProbes-Placenta_MeanB_allProbes >= 0.2)


# for top 2% probes, delta beta
Placenta_MeanB <- NormBAll_placenta[rownames(NormBAll_placenta)%in%ProbeContrib.df.orderedTop2perc$Name, 
                                    colnames(NormBAll_placenta)%in%Placenta_names] %>% 
  rowMeans()
Placenta_MeanB <- Placenta_MeanB[match(ProbeContrib.df.orderedTop2perc$Name, names(Placenta_MeanB))]

Outlier_MeanB <- NormBAll_placenta[rownames(NormBAll_placenta)%in%ProbeContrib.df.orderedTop2perc$Name, 
                                      colnames(NormBAll_placenta)%in%Outlier_names] %>% 
  rowMeans()
Outlier_MeanB <- Outlier_MeanB[match(ProbeContrib.df.orderedTop2perc$Name, names(Outlier_MeanB))]
 

table(abs(Outlier_MeanB-Placenta_MeanB) >= 0.2)

table(names(Placenta_MeanB)==ProbeContrib.df.orderedTop2perc$Name)
which(abs(Outlier_MeanB-Placenta_MeanB) >= 0.2) %>% is.unsorted()

table(Outlier_MeanB-Placenta_MeanB >= 0.2)

```

Are these top 2% probes from PC1 located in PMDs?

How many probes in PMDs and ICRs:
```{r}
## Get probe ranges and select probe not related with islands
Gset_placenta <- makeGenomicRatioSetFromMatrix(NormBMlist$NormBAll, array = "IlluminaHumanMethylation450k",
                                               annotation = "ilmn12.hg19",mergeManifest = TRUE,
                                               what = "Beta")

GRangeProbe <- Gset_placenta@rowRanges[, c(1:2,7)]

# add mcol to GRangeProbe
## get GRange of all probes: 200k
### add values to GRange object metadata columns
matchWithGRprobe <- base::match(GRangeProbe$Name, rownames(NormBMlist$NormBAll))
values(GRangeProbe) <- cbind(values(GRangeProbe), DataFrame(NormBMlist$NormBAll[matchWithGRprobe,], check.names = F)) 

### add seqinfo
bs_hg19 <- BSgenome.Hsapiens.UCSC.hg19
### check names are the same or not
seqnames(seqinfo(bs_hg19))[1:24] == seqinfo(GRangeProbe) %>% seqnames()
### assign seqinfo to my GRange for all probes 626374
seqinfo(GRangeProbe) <- Seqinfo(seqnames = seqnames(seqinfo(bs_hg19))[1:24],
                                seqlengths = seqlengths(seqinfo(bs_hg19))[1:24],
                                isCircular = isCircular(seqinfo(bs_hg19))[1:24],
                                genome = genome(seqinfo(bs_hg19))[1:24])

plyranges::find_overlaps(PMDs_hg19, GRangeProbe[,1:3])$Name %>% length() 

plyranges::find_overlaps(ICRsGR, GRangeProbe[,1:3])$Name %>% length() 

```
* In total, there were 44353 probes from 300k common probes (408 samples)

How many top 2% probes in PMDs and ICRs:
```{r}
Top2GRangeProbe <- GRangeProbe[,1:3] %>% plyranges::filter(Name%in%ProbeContrib.df.orderedTop2perc$Name)

plyranges::find_overlaps(PMDs_hg19, Top2GRangeProbe)$Name %>% length() 

plyranges::find_overlaps(ICRsGR, Top2GRangeProbe)$Name %>% length() 


# placenta specific ICRs
ICRsGR_placentaSpe <- ICRsGR[ICRsGR$ICR_Type%in%"PlacentaSpecific"]

plyranges::find_overlaps(ICRsGR_placentaSpe, Top2GRangeProbe)$Name %>% length() 

```
* There are 2018 probes from top 2 percent probes (6016) from PC1 that located in PMDs.


# T1 placenta 
```{r T1}

PhenodataPlacenta_T1 <- Matched_pd[Matched_pd$Sample%in%"Placenta" & Matched_pd$Trimester%in%"First",]

Placenta_Mnorm_v1_T1 <-  NormBMlist$NormMAll[,colnames(NormBMlist$NormMAll)%in%PhenodataPlacenta_T1$Sample_name]
Placenta_betaNorm_T1 <- 2^Placenta_Mnorm_v1_T1 / (1+2^Placenta_Mnorm_v1_T1)

```


# T1 outlier and pure samples 
```{r}
## plot box plot for beta values of Dim1_top2perc porbes
## manipulate data frames #"GSM1702177_7970368142_R05C02"
# PhenodataPlacenta_T1$Sample_name[match(CutOffRightNames, PhenodataPlacenta_T1$Sample_name)]

OutlierT1 <- tibble(Outlier_T1Names = CutOffRightNames[CutOffRightNames%in%PhenodataPlacenta_T1$Sample_name],
                    Outlier_T1 = paste0("Outlier_T1_", 1:5))

PhenodataPlacenta_T1_addOutlier <- PhenodataPlacenta_T1 %>% left_join(OutlierT1, by=c("Sample_name"="Outlier_T1Names"))

PhenodataPlacenta_T1_addOutlier$Outlier_T1[is.na(PhenodataPlacenta_T1_addOutlier$Outlier_T1)] <- "Pure"

phenoDataPlacenta.df_T1 <- PhenodataPlacenta_T1_addOutlier %>% as.data.frame(stringsAsFactors=FALSE) %>%
  dplyr::select(c("Sample_name", "Outlier_T1"))

```

# plot box plot for mixed and pure placenta samples:
```{r}

placenta_beta_dim1Top2perc.df_T1 <- Placenta_betaNorm_T1[rownames(Placenta_betaNorm_T1)%in%ProbeContrib.df.orderedTop2perc$Name,] %>% 
  as.data.frame() %>% 
  rownames_to_column(var="Name") %>% 
  melt(id.var="Name") %>% 
  `colnames<-`(c("Name", "Sample_name", "value")) %>% 
  left_join(phenoDataPlacenta.df_T1, by="Sample_name")

outlierDim1Plot_T1 <- placenta_beta_dim1Top2perc.df_T1 %>% 
  ggplot(aes(x=Outlier_T1, y=value))+
  geom_boxplot()+
  # scale_x_discrete(labels=c("Outlier"="Outliers", "Standard"="Other Samples"))+
  labs(x="", y="Beta values")+
  # stat_compare_means(label = "p.signif", method = "t.test",
  #                    ref.group = ".all.") +
  theme(axis.text.y = element_text(size=9, colour ="black"),
        axis.text.x = element_text(size=9, colour ="black"),
        axis.title=element_text(size=9, colour ="black"),#,face="bold"
        text = element_text(size=9, colour ="black"))

outlierDim1Plot_T1

tiff(file = "/home/qianhui/DNAme/Process_decidual/figures/Placenta_Dim1top2percProbePlot_ForOutlier_T1.tiff",
     width = 20, height = 9, units = "cm", res = 300)
theme_set(theme_bw())
outlierDim1Plot_T1
dev.off()

# quantile(PCA_Placenta_M_T2$ind$coord[, 1], 0.75)+ 1.5*IQR(PCA_Placenta_M_T2$ind$coord[, 1])
# quantile(PCA_Placenta_M_T2$ind$coord[, 1], 0.25)- 1.5*IQR(PCA_Placenta_M_T2$ind$coord[, 1])

saveRDS(placenta_beta_dim1Top2perc.df_T1, file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/placenta_beta_dim1Top2perc.df_T1.rds")

```



# T2 placenta, and pot beta values for outliers and pure samples
```{r T2}

PhenodataPlacenta_T2 <- Matched_pd[Matched_pd$Sample%in%"Placenta" & Matched_pd$Trimester%in%"Second",]

Placenta_Mnorm_v1_T2 <-  NormBMlist$NormMAll[,colnames(NormBMlist$NormMAll)%in%PhenodataPlacenta_T2$Sample_name]
Placenta_betaNorm_T2 <- 2^Placenta_Mnorm_v1_T2 / (1+2^Placenta_Mnorm_v1_T2)

```


# T2 outlier and pure samples 
```{r}
## plot box plot for beta values of Dim1_top2perc porbes
## manipulate data frames #"GSM1702177_7970368142_R05C02"
PhenodataPlacenta_T2$Outlier <- factor(ifelse(PhenodataPlacenta_T2$Sample_name%in%CutOffRightNames, 
                                              "Outlier_T2_1", "Pure"))


phenoDataPlacenta.df_T2 <- PhenodataPlacenta_T2 %>% as.data.frame(stringsAsFactors=FALSE) %>%
  dplyr::select(c("Sample_name", "Outlier"))

placenta_beta_dim1Top2perc_T2 <- Placenta_betaNorm_T2[rownames(Placenta_betaNorm_T2)%in%ProbeContrib.df.orderedTop2perc$Name,]
placenta_beta_dim1Top2perc.df_T2 <- placenta_beta_dim1Top2perc_T2 %>% 
  as.data.frame() %>% rownames_to_column(var="Name") %>% melt(id.var="Name") %>% 
  `colnames<-`(c("Name", "Sample_name", "value")) %>% 
  left_join(phenoDataPlacenta.df_T2, by="Sample_name")

outlierDim1Plot_T2 <- placenta_beta_dim1Top2perc.df_T2 %>% 
  ggplot(aes(x=Outlier, y=value))+
  geom_boxplot()+
  # scale_x_discrete(labels=c("Outlier"="Outliers", "Standard"="Other Samples"))+
  labs(x="", y="Beta values")+
  # stat_compare_means(label = "p.signif", method = "t.test",
  #                    ref.group = ".all.") +
  theme(axis.text.y = element_text(size=12, colour ="black"),
        axis.text.x = element_text(size=12, colour ="black"),
        axis.title=element_text(size=12, colour ="black"),#,face="bold"
        text = element_text(size=12, colour ="black"))

outlierDim1Plot_T2

tiff(file = "/home/qianhui/DNAme/Process_decidual/figures/Placenta_Dim1top2percProbePlot_ForOutlier_T2_residual.tiff",
     width = 20, height = 12, units = "cm", res = 300)
theme_set(theme_bw())
outlierDim1Plot_T2
dev.off()

# quantile(PCA_Placenta_M_T2$ind$coord[, 1], 0.75)+ 1.5*IQR(PCA_Placenta_M_T2$ind$coord[, 1])
# quantile(PCA_Placenta_M_T2$ind$coord[, 1], 0.25)- 1.5*IQR(PCA_Placenta_M_T2$ind$coord[, 1])

saveRDS(placenta_beta_dim1Top2perc.df_T2, file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/placenta_beta_dim1Top2perc.df_T2.rds")

```

* The results showed that this one outlier from second trimester could be an impure placenta sample.

# Term placenta 
```{r T1}

PhenodataPlacenta_Term <- Matched_pd[Matched_pd$Sample%in%"Placenta" & Matched_pd$Trimester%in%"Term",]

Placenta_Mnorm_v1_Term <-  NormBMlist$NormMAll[,colnames(NormBMlist$NormMAll)%in%PhenodataPlacenta_Term$Sample_name]
Placenta_betaNorm_Term <- 2^Placenta_Mnorm_v1_Term / (1+2^Placenta_Mnorm_v1_Term)

```



# Term outlier and pure samples 
```{r}
## plot box plot for beta values of Dim1_top2perc porbes
## manipulate data frames #"GSM1702177_7970368142_R05C02"

OutlierTerm <- tibble(
  Outlier_TermNames = CutOffRightNames[CutOffRightNames%in%PhenodataPlacenta_Term$Sample_name],
  Outlier_Term = paste0("Outlier_Term_", 1:5))

PhenodataPlacenta_Term_addOutlier <- PhenodataPlacenta_Term %>% left_join(OutlierTerm, by=c("Sample_name"="Outlier_TermNames"))

PhenodataPlacenta_Term_addOutlier$Outlier_Term[is.na(PhenodataPlacenta_Term_addOutlier$Outlier_Term)] <- "Pure"

# make df for plotting
phenoDataPlacenta.df_Term <- PhenodataPlacenta_Term_addOutlier %>% as.data.frame(stringsAsFactors=FALSE) %>%
  dplyr::select(c("Sample_name", "Outlier_Term"))

placenta_beta_dim1Top2perc_Term <- Placenta_betaNorm_Term[rownames(Placenta_betaNorm_Term)%in%ProbeContrib.df.orderedTop2perc$Name,]

placenta_beta_dim1Top2perc.df_Term <- placenta_beta_dim1Top2perc_Term %>% 
  as.data.frame() %>% rownames_to_column(var="Name") %>% melt(id.var="Name") %>% 
  `colnames<-`(c("Name", "Sample_name", "value")) %>% 
  left_join(phenoDataPlacenta.df_Term, by="Sample_name")

outlierDim1Plot_Term <- placenta_beta_dim1Top2perc.df_Term %>% 
  ggplot(aes(x=Outlier_Term, y=value))+
  geom_boxplot()+
  # scale_x_discrete(labels=c("Outlier"="Outliers", "Standard"="Other Samples"))+
  labs(x="", y="Beta values")+
  # stat_compare_means(label = "p.signif", method = "t.test",
  #                    ref.group = ".all.") +
  theme(axis.text.y = element_text(size=12, colour ="black"),
        axis.text.x = element_text(size=12, colour ="black"),
        axis.title=element_text(size=12, colour ="black"),#,face="bold"
        text = element_text(size=12, colour ="black"))

outlierDim1Plot_Term

tiff(file = "/home/qianhui/DNAme/Process_decidual/figures/Placenta_Dim1top2percProbePlot_ForOutlierTerm.tiff",
     width = 20, height = 12, units = "cm", res = 300)
theme_set(theme_bw())
outlierDim1Plot_Term
dev.off()

# quantile(PCA_Placenta_M_T2$ind$coord[, 1], 0.75)+ 1.5*IQR(PCA_Placenta_M_T2$ind$coord[, 1])
# quantile(PCA_Placenta_M_T2$ind$coord[, 1], 0.25)- 1.5*IQR(PCA_Placenta_M_T2$ind$coord[, 1])

saveRDS(placenta_beta_dim1Top2perc.df_Term, file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/placenta_beta_dim1Top2perc.df_Term.rds")

```


# arrange plots

## combine top 2% boxplot to 1 plot
```{r}
#placenta_beta_dim1Top2perc.df_T1
# placenta_beta_dim1Top2perc.df <- placenta_beta_dim1Top2perc.df %>% dplyr::select(-OutlierAdd2)
colnames(placenta_beta_dim1Top2perc.df_T1)[4] <- "Outlier"
colnames(placenta_beta_dim1Top2perc.df_T2)[4] <- "Outlier"
colnames(placenta_beta_dim1Top2perc.df_Term)[4] <- "Outlier"

placenta_beta_dim1Top2perc.df_T1$Trimester <- rep("First", nrow(placenta_beta_dim1Top2perc.df_T1))

placenta_beta_dim1Top2perc.df_T2$Trimester <- rep("Second", nrow(placenta_beta_dim1Top2perc.df_T2))

placenta_beta_dim1Top2perc.df_Term$Trimester <- rep("Term", nrow(placenta_beta_dim1Top2perc.df_Term))

placenta_beta_dim1Top2perc.df_all <- rbind(placenta_beta_dim1Top2perc.df_T1, 
                                           placenta_beta_dim1Top2perc.df_T2,
                                           placenta_beta_dim1Top2perc.df_Term)

placenta_beta_dim1Top2perc.df_all$OutlierColor <- str_replace_all(placenta_beta_dim1Top2perc.df_all$Outlier,
                                                                  "Outlier.+", "Outlier")


saveRDS(placenta_beta_dim1Top2perc.df_all, file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/placenta_beta_dim1Top2perc.df_all.rds")

theme_set(theme_pubr())

outlierDim1Plot_all <- placenta_beta_dim1Top2perc.df_all %>% 
  ggplot(aes(x=as.factor(Outlier), y=value, fill=OutlierColor))+ 
  geom_violin(trim = T)+
  geom_boxplot(width=0.1, fill="white")+
  scale_fill_manual(name="Placenta", 
                    values = c("Outlier"="#1B9E77", "Pure"="#D95F02"),
                    labels = c("Outlier"="Mixed", "Pure"="Pure"))+
  # scale_x_discrete(labels="")+
  labs(x="", y="Beta values")+ #, fill="Placenta"
  facet_wrap(~Trimester, scale="free_x", nrow = 2)+
  # facet_grid(.~Trimester, scale="free_x")+
  guides(fill=guide_legend(nrow = 1)) +
  theme(axis.text.y = element_text(size=9, colour ="black"),
        axis.text.x = element_text(size=9, colour ="black"),
        # axis.ticks.x= element_blank(),
        axis.title=element_text(size=9, colour ="black"),#,face="bold"
        strip.text.x = element_text(size = 9, colour = "black"),
        text = element_text(size=9, colour ="black"),
        legend.position="top")

# convert ggplot object to grob object
gp <- ggplotGrob(outlierDim1Plot_all)

# optional: take a look at the grob object's layout
gtable::gtable_show_layout(gp)

# get gtable columns corresponding to the facets (5 & 9, in this case)
facet.columns <- gp$layout$l[grepl("panel", gp$layout$name)]

# get the number of unique x-axis values per facet (1 & 3, in this case)
x.var <- sapply(ggplot_build(outlierDim1Plot_all)$layout$panel_scales_x,
                function(l) length(l$range$range))

# change the relative widths of the facet columns based on
# how many unique x-axis values are in each facet
gp$widths[facet.columns] <- gp$widths[facet.columns] * c(2,6,2)

# plot result
grid::grid.draw(gp)

saveRDS(gp, file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/outlierDim1Plot_all.rds")
# saveRDS(placenta_beta_dim1Top2perc.df_all, file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/placenta_beta_dim1Top2perc.df_all.rds")

```


# Figure1
```{r}

theme_set(theme_pubr())

# PCAs
# PCA_Placenta_M, PCA_PlacentaT1_M, PCA_PlacentaTerm_M
# load(file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/PCA_placenta_deciduals_allT1Term_BMval.RData")

PCA408plot <- readRDS(file = "/home/qianhui/DNAme/Process_decidual/figures/T1TermSamples_PCA408samples.rds")

# PCAT1T2Termplot <- readRDS(file = "/home/qianhui/DNAme/Process_decidual/figures/PCAallplot_noSexLabel.rds")

PCAT1T2Termplot <- readRDS(file = "/home/qianhui/DNAme/Process_decidual/figures/PCAallplot.rds")

outlierDim1Plot_all <- readRDS(file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/outlierDim1Plot_all.rds")

# top2perc_GOPlot <- readRDS(file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/top2perc_GOPlot.rds")

# PCAs <- ggarrange(PCA551plot, 
#                   PCAT1plot, PCAT2plot, PCATermplot,
#                   labels = c("A", "B", "C", "D"),
#                   ncol = 4, nrow = 1, widths = c(1,1,1,1.1),
#                   font.label=list(size = 12, color = "black", face = "bold", family = "Arial"),
#                   legend=FALSE)

PCAs <- ggarrange(PCA408plot, 
                  # PCAT1plot, PCAT2plot, PCATermplot,
                  PCAT1T2Termplot,
                  labels = c("A", "B"),
                  ncol = 2, nrow = 1, widths = c(1.3,3),
                  font.label=list(size = 12, color = "black", face = "bold", family = "Arial"),
                  legend=FALSE)

# PC1plots <- ggarrange(outlierDim1Plot_all, #top2perc_GOPlot, 
#           labels = c("C", "D"),
#           font.label = list(size = 12, face = "bold", color ="black"),
#           label.x = c(0, 0.05),
#           ncol = 2, widths = c(1,0.8))

# tiff(file = "/home/qianhui/DNAme/Process_decidual/figures/TechFigure1_v4_noSexLabel.tiff",
#      width = 25, height = 17, units = "cm", res = 300)
tiff(file = "/home/qianhui/DNAme/Process_decidual/figures/TechFigure1_v4.tiff",
     width = 25, height = 17, units = "cm", res = 300)
theme_set(theme_pubr())
ggarrange(
               PCAs,
               # PCA551plot,PCAT1plot, PCATermplot,
               outlierDim1Plot_all,
               # outlierDim1Plot_T1Term, top2perc_GOPlot_T1Term, 
               # densityCorBoxplots,
               labels = c("", "C"),
               # ncol = 3, 
               nrow = 2,
               heights = c(1.2,2),
               
               # widths = c(1,1,1, 1,2,1),
               font.label = list(size = 12, face = "bold", color ="black")
               # legend="right",
               # common.legend = TRUE
               # layout_matrix = rbind(c(1,2,NA),c(3,3,3), c(4,5,6))
               )
dev.off()


```


# check whether 2% probes have difference between fetal sex group and trimester group, suplimentary file

Add fetal sex column and Study column into data frame `placenta_beta_dim1Top2perc.df_all`

```{r}

placenta_beta_dim1Top2perc.df_all <- placenta_beta_dim1Top2perc.df_all %>% 
  left_join(Matched_pd_placenta[,c("Sample_name", "Fetal_Sex", "ArrayType", "Study", "Outlier")], by="Sample_name")

# saveRDS(placenta_beta_dim1Top2perc.df_all, file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/placenta_beta_dim1Top2perc.df_all.rds")

```


Test whether top 2% probes is different between trimester groups, fetal sex groups and batch groups:

2% probes:
```{r}

TrimesterTop2 <- placenta_beta_dim1Top2perc.df_all %>%
  ggplot(aes(x=Trimester, y=value)) +
  geom_boxplot()+
  labs(x="", y="Beta values")+
  # stat_compare_means(label = "p.signif", method = "wilcox.test",
  #                    ref.group = ".all.") +
  theme(axis.text.y = element_text(size=9, colour ="black"),
        axis.text.x = element_text(size=9, colour ="black"),
        axis.title=element_text(size=9, colour ="black"),#,face="bold"
        text = element_text(size=9, colour ="black"))


SexTop2 <- placenta_beta_dim1Top2perc.df_all %>%
  ggplot(aes(x=Fetal_Sex, y=value)) +
  geom_boxplot()+
  labs(x="", y="Beta values")+
  # stat_compare_means(label = "p.signif", method = "wilcox.test",
  #                    ref.group = ".all.") +
  theme(axis.text.y = element_text(size=9, colour ="black"),
        axis.text.x = element_text(size=9, colour ="black"),
        axis.title=element_text(size=9, colour ="black"),#,face="bold"
        text = element_text(size=9, colour ="black"))


StudyTop2 <- placenta_beta_dim1Top2perc.df_all %>%
  ggplot(aes(x=Study, y=value)) +
  geom_boxplot()+
  labs(x="", y="Beta values")+
  # stat_compare_means(label = "p.signif", method = "wilcox.test",
  #                    ref.group = ".all.") +
  theme(axis.text.y = element_text(size=9, colour ="black"),
        axis.text.x = element_text(size=9, colour ="black", angle = 45, hjust = 1),
        axis.title=element_text(size=9, colour ="black"),#,face="bold"
        text = element_text(size=9, colour ="black"))

OutlierTop2 <- placenta_beta_dim1Top2perc.df_all %>%
  ggplot(aes(x=Outlier.y, y=value)) +
  geom_boxplot()+
  labs(x="", y="Beta values")+
  # stat_compare_means(label = "p.signif", method = "wilcox.test",
  #                    ref.group = ".all.") +
  theme(axis.text.y = element_text(size=9, colour ="black"),
        axis.text.x = element_text(size=9, colour ="black"),
        axis.title=element_text(size=9, colour ="black"),#,face="bold"
        text = element_text(size=9, colour ="black"))

# library("ggpubr")
# ggline(placenta_beta_dim1Top2perc.df_all, x = "Trimester", y = "value", 
#        add = c("mean_se"), 
#        order = c("First", "Second", "Term"),
#        ylab = "Beta", xlab = "Trimester")+
#   # ylim(0,1)
#   scale_y_continuous(breaks=seq(0.0, 1, 0.2), limits=c(0, 1))+
#   geom_hline(yintercept = c(0.35,0.5))

tiff(file = "/home/qianhui/DNAme/Process_decidual/figures/TechSupFigure3.tiff",
     width = 25, height = 10, units = "cm", res = 300)
theme_set(theme_pubr())

ggarrange(TrimesterTop2, SexTop2, StudyTop2, OutlierTop2)

dev.off()

```



```{r}

TestPCFUN <- function(PC, pd){
  
  mod1 = lm(PC ~ pd)
  anova_res <- anova(mod1)
  data.frame(Fvalue=anova_res$`F value`[1], Pvalue=anova_res$`Pr(>F)`[1])
}
  
```


PC1/2 and Trimester

```{r}

PC1_trimester <- TestPCFUN(PC = PCA_Placenta_M$ind$coord[,1], pd = Matched_pd$Trimester) %>% set_rownames("PC1_trimester")


PC2_trimester <- TestPCFUN(PC = PCA_Placenta_M$ind$coord[,2], pd = Matched_pd$Trimester) %>% set_rownames("PC2_trimester")


```

PC1/2 and sex:
```{r}

PC1_sex <- TestPCFUN(PC = PCA_Placenta_M$ind$coord[,1], pd = Matched_pd$Fetal_Sex) %>% set_rownames("PC1_sex")


PC2_sex <- TestPCFUN(PC = PCA_Placenta_M$ind$coord[,2], pd = Matched_pd$Fetal_Sex) %>% set_rownames("PC2_sex")


```

PC1/2 and batch (Study):
```{r}

PC1_study <- TestPCFUN(PC = PCA_Placenta_M$ind$coord[,1], pd = Matched_pd$Study) %>% set_rownames("PC1_study")


PC2_study <- TestPCFUN(PC = PCA_Placenta_M$ind$coord[,2], pd = Matched_pd$Study) %>% set_rownames("PC2_study")


```

Pc1/2 and samples
```{r}

PC1_tissue <- TestPCFUN(PC = PCA_Placenta_M$ind$coord[,1], pd = Matched_pd$Sample) %>% set_rownames("PC1_tissue")


PC2_tissue <- TestPCFUN(PC = PCA_Placenta_M$ind$coord[,2], pd = Matched_pd$Sample) %>% set_rownames("PC2_tissue")

```


bind all df and write csv:
```{r}

rbind(PC1_trimester, PC2_trimester,
      PC1_sex, PC2_sex,
      PC1_study, PC2_study,
      PC1_tissue, PC2_tissue) %>% 
  rownames_to_column(var="compare") %>% 
  arrange(-Fvalue) %>% 
  write_csv(path = "/home/qianhui/DNAme/Process_decidual/RDSfiles/PC1PC2_TrimesterSexBatch.csv")

```


