---
title: "technical report (mainly 450k data, 15 GEO data sets)"
author: "Qianhui"
date: "28/08/2018"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", error = TRUE, echo = TRUE) # fig.width = 7, 
```


# 1.Packages and data required

```{r package, warning=FALSE, message=FALSE}
# library(TxDb.Hsapiens.UCSC.hg19.knownGene)
# library(BSgenome.Hsapiens.UCSC.hg19)

library(dplyr)
library(tibble)
library(readr)
library(magrittr)
library(data.table)
library(tibble)
library(stringi)
library(tidyverse)
library(reshape2)
library(stringr)
library(minfi)
library(IlluminaHumanMethylation450kmanifest)
library(IlluminaHumanMethylation450kanno.ilmn12.hg19)
library(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
library(matrixStats)
library(wateRmelon)
library(limma)
library(missMethyl)
# library(ewastools)
library(ENmix)
library(ChAMP)
library(FactoMineR)
library(plyranges)

library(GenomicRanges)
library(RColorBrewer)
library(ggplot2)
library(plotrix)
library(gridExtra)
library(doParallel)

library(GEOquery)
# colours to use
library(RColorBrewer)
pal <- brewer.pal(8, "Dark2")


colorFun <- colorRampPalette(brewer.pal(9, "Set1"))
pal2 <- colorFun(10)

theme_set(theme_bw())

```


# 2. get phenodata/metadata for all GEO data sets (15)
'GSE41933'--unbilical cord; 'GSE66459'--unbilical cord blood

```{r meta, eval=FALSE}
# Download of the meta data for each data set
# GSE73375 <- getGEO('GSE73375', destdir  = '/home/qianhui/DNAme/GAclock/RepeatBensClock/GEO_data/')

GEO_num <- c('GSE66210', 'GSE74738', 'GSE69502', 'GSE75196', 
             'GSE75248', 'GSE93208', 'GSE115508', 'GSE120250', 
             'GSE108567', 'GSE100197', 'GSE98224', 'GSE106089', 
             'GSE71678', 'GSE113600', 'GSE66459')

#GSE93429, GSE79262, GSE41933 removed because they have no raw IDAT files.
#GSE108567 removed because it's a repeated data set for reanalysis purpose.

library(doParallel)
registerDoParallel(cores = 10)

for(i in 1:15){
  assign(paste0(GEO_num[i]),
         getGEO(GEO_num[i], destdir  = '/home/qianhui/DNAme/Process_decidual/GEO_dataSets/MetaData_GEO15'))
}

# Extract the meta data
# GSE73375.meta <- pData(phenoData(GSE73375[[1]]))

for(i in 1:15){
  # if data sets contain data form differenct platforms
  if(length(get(GEO_num[i]))>1){
    
  DNAmePlatform <- str_detect(names(get(GEO_num[i])), pattern = 'GPL13534')
  assign(paste0(GEO_num[i], ".meta"),
         pData(phenoData(get(GEO_num[i])[DNAmePlatform][[1]])))
  
  }else{
  # if data only from one platform, we can extract meta data directly 
  assign(paste0(GEO_num[i], ".meta"),
         pData(phenoData(get(GEO_num[i])[[1]])))
  } # end if
} # end for

# Save all the meta data into one file
save(GSE66210.meta, GSE74738.meta, GSE69502.meta, GSE75196.meta, GSE75248.meta, 
     GSE93208.meta, GSE115508.meta, GSE120250.meta, GSE108567.meta, GSE100197.meta, 
     GSE98224.meta, GSE106089.meta, GSE71678.meta, GSE113600.meta, GSE66459.meta,
     file='/home/qianhui/DNAme/Process_decidual/RDSfiles/pdataOrMetadata_GEO_15datasets.Rdata' )

```


# 3. get raw data
save raw data to the baseDir

```{r, eval=FALSE}

registerDoParallel(cores = 10)

for(i in 1:15){
  # assign(paste0(GEO_num[i], ".raw"),
  getGEOSuppFiles(GEO_num[i], 
                  # makeDirectory = FALSE,
                  # filter_regex='RAW',
                  baseDir = '/home/qianhui/DNAme/Process_decidual/GEO_dataSets')
}

```


# 4. untar and unzip raw data (IDAT files)
First untar .tar files in bash, then unzip files in R
```{r, eval=FALSE}

GEO_num %>% head()

write.table(GEO_num, file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_GSEids.txt", 
            quote = FALSE, 
            row.names = FALSE, col.names = FALSE)

for(n in 1:15){
# GEOid <- GEO_num[n]
  dir <- '/home/qianhui/DNAme/Process_decidual/GEO_dataSets/'
  gzFileNames <- list.files(paste0('/home/qianhui/DNAme/Process_decidual/GEO_dataSets/',
                                   GEO_num[n]), pattern = ".gz")
# some folder have more than one gz files, so I used if else function
if(length(gzFileNames)==0){
  print("no gz files, no need to use gunzip")

  }else{

  for(gz in gzFileNames){
  gunzip(filename = paste0(dir, GEO_num[n], "/", gz), overwrite = TRUE)
  } # end for gz
 } # end if
} # end for n

```
😀


# 5.reorganise meta data for all GEO data sets

Read the data to RG channel Set using `minfi` package.

## function to get RGsets:

```{r}
##getwd()

ann450k <- getAnnotation(IlluminaHumanMethylation450kanno.ilmn12.hg19)
saveRDS(ann450k, file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/ann450k_GEO15sets.rds")

annEPIC <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
saveRDS(annEPIC, file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/annEPIC_GEO15sets.rds")
# get RGset for GSE66210
# GSE66210.RGset <- getRGset('GSE66210')

```

## Manipulate meta data for GEO datasets:
```{r ReadData1}
source(file = "~/DNAme/Process_decidual/Rcodes/TrimGEOmetaDataFUN.R")

TrimGEOmetaDataFUN(GSEnum = "all")

```

## our data
### Add our outliers 5 + 5 normal placenta, `GEO_allMetaData_v1.RData`
```{r ReadOurOrYourData}

# load(file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_allMetaData_v1.RData")

load(file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_allMetaData_v1_all.RData")


OurOrYourDataPhenoFUN <- function(EPICdata){

  if(EPICdata=="OurEPIC10"){
baseDirOurEPIC <- file.path("/home/qianhui/DNAme/EPICarrayData/idat_files")
phenotypes <- read.metharray.sheet(baseDirOurEPIC)

set.seed(2018)
phenotypes_GA6to9standard <- phenotypes %>% 
  dplyr::filter(Pool_ID%in%"Placental Tissue") %>% 
  dplyr::filter(Sample_Group%in%c(6:9)) %>% 
  dplyr::filter(Outliers%in%"Standard") %>% 
  sample_n(5)

phenotypes_GA6to9_10samples_inGEOformat <-  phenotypes %>% 
  dplyr::filter(Pool_ID%in%"Placental Tissue") %>% 
  dplyr::filter(Sample_Group%in%c(6:9)) %>% 
  dplyr::filter(Outliers%in%"Outlier") %>% 
  bind_rows(phenotypes_GA6to9standard) %>% 
  inset('GEO_accession', value=rep("OurData", 10)) %>% 
  inset('Sample', value=rep('Placenta', 10)) %>% 
  mutate(Sample_name = str_replace_all(Basename, ".+/", "")) %>% 
  inset('Disease', value=rep("Uncomplicated", 10)) %>% 
  mutate(Gestation= as.character(Gestational.Age)) %>% 
  mutate(Trimester= str_replace_all(Trimester, "1st-Trimester", "First")) %>% 
  mutate(Additional_Info =Outliers) %>% 
  inset('Additional_Info_2', value=rep("NA", 10)) %>% 
  mutate(Fetal_Sex=str_replace_all(Fetal.Sex, c('M'='Male', 'F'='Female'))) %>% 
  inset('ArrayType', value=rep('EPIC',10)) %>% 
  inset('Study', value=rep('ourStudy', 10)) %>% 
  dplyr::select(GEO_accession:Gestation, Trimester, Additional_Info:Study, Array, Slide)

# I want to copy these 10 samples to IDATall folder, so I need to write this txt first 
write.table(phenotypes_GA6to9_10samples_inGEOformat$Sample_name, 
            file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/ourStudy10samples.txt", 
            quote = FALSE, 
            row.names = FALSE, col.names = FALSE)

# bind our phenodata with GEO meta data
save(GSE66210.meta_v1, GSE74738.meta_v1, GSE69502.meta_v1, GSE75196.meta_v1, 
     GSE75248.meta_v1, GSE93208.meta_v1, GSE115508.meta_v1, GSE120250.meta_v1, 
     GSE100197.meta_v1, GSE98224.meta_v1, GSE106089.meta_v1, GSE71678.meta_v1, 
     GSE113600.meta_v1, GSE66459.meta_v1, phenotypes_GA6to9_10samples_inGEOformat,
     file = paste0("/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_allMetaData_v1_add", EPICdata, ".RData"))
  }
  
}

```


### use `OurOrYourDataPhenoFUN` to add phenodata of our data sets to `GEO_allMetaData_v1.RData`
```{r}

EPICdata <- "OurEPIC10"

OurOrYourDataPhenoFUN(EPICdata="OurEPIC10")

```


# 6. bind all the meta data together and read `GEOmeta` samplesheet. 
  Then read the data we wanted to Red-Green channel sets.
```{r}

# load(file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_allMetaData_v1.RData")
load(file = paste0("/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_allMetaData_v1_add", EPICdata, ".RData"))


GEOmeta <- rbind(GSE66210.meta_v1, GSE74738.meta_v1, GSE69502.meta_v1, GSE75196.meta_v1, 
                 GSE75248.meta_v1, GSE93208.meta_v1, GSE115508.meta_v1, GSE120250.meta_v1, 
                 GSE100197.meta_v1, GSE98224.meta_v1, GSE106089.meta_v1, GSE71678.meta_v1, 
                 GSE113600.meta_v1, GSE66459.meta_v1, phenotypes_GA6to9_10samples_inGEOformat)

# remove the samples with no trimester informations
GEOmeta <- GEOmeta[!(GEOmeta$Trimester%in%"NA"), ]

write_csv(GEOmeta, path = "/home/qianhui/DNAme/Process_decidual/GEO_dataSets/IDAT_all/GEOmeta.csv")

GEO_baseDir <- file.path("/home/qianhui/DNAme/Process_decidual/GEO_dataSets/IDAT_all")

GEO_phenotypes <- read.metharray.sheet(GEO_baseDir, pattern = "GEOmeta.csv$")

save(GEOmeta, GEO_phenotypes, 
     file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_phenotypes_add10.RData")

```


## read uncomplicated placental sampels 

We read 450K and EPIC data separately, because they have different annotations.

### load phenodata, start from here, so we don't need to redo all steps above.
```{r}

load(file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_phenotypes_add10.RData")

# filter out some data sets /samples that we don't need
## remove complicated samples by Disease column (-864)
GEO_phenotypes_uncom <- GEO_phenotypes[GEO_phenotypes$Disease%in%"Uncomplicated",]

## remove samples that were replicate samples in GSE66210 (-4)
Du <- c("GSM1617040","GSM1617041", "GSM1617042","GSM1617043")
GEO_phenotypes_mGSE66210 <- GEO_phenotypes_uncom[!(GEO_phenotypes_uncom$Study%in%"GSE66210" & GEO_phenotypes_uncom$GEO_accession%in%Du),]

## remove perterm samples from GSE74738 (-15)
GEO_phenotypes_mGSE74738 <- GEO_phenotypes_mGSE66210 %>% 
  dplyr::filter(!(Study%in%"GSE74738" & Sample%in%c("placental mesenchyme", "placental trophoblast"))) %>% 
  dplyr::filter(!(Study%in%"GSE74738" & Trimester%in%"Third"))
  
## remove samples from GSE93208 (these samples were isolated cells not bulk tissue) (-19)
## and GSE106089 (these samples were all preterm) (-46), so -65 intotal.
GEO_phenotypes_mGSE93208 <- GEO_phenotypes_mGSE74738 %>% 
  dplyr::filter(!(GEO_phenotypes_mGSE74738$Study%in%c("GSE93208","GSE106089")))

## remove preterm samples from GSE98224 (-9)
GEO_phenotypes_mGSE98224 <- GEO_phenotypes_mGSE93208 %>% 
  dplyr::filter(!(Study%in%"GSE98224" & Trimester%in%"Third"))

## remove preterm samples from GSE71678 (-4)
GEO_phenotypes_mGSE71678 <- GEO_phenotypes_mGSE98224 %>% 
  dplyr::filter(!(Study%in%"GSE71678" & Trimester%in%"Third"))

## remove preterm samples from GSE66459 (-11)
GEO_phenotypes_mGSE66459 <- GEO_phenotypes_mGSE71678 %>% 
  dplyr::filter(!(Study%in%"GSE66459" & Trimester%in%"Third"))

## remove preterm samples from GSE115508 (-35)
GEO_phenotypes_mGSE115508 <- GEO_phenotypes_mGSE66459 %>% 
  dplyr::filter(!(Study%in%"GSE115508" & Trimester%in%"Third"))

## remove samples that were not placenta in GSE69502 (-49)
GEO_phenotypes_mGSE69502 <- GEO_phenotypes_mGSE115508 %>% 
  dplyr::filter(!(Study%in%"GSE69502" & !(Sample%in%"Placenta")))

GEO_phenotypes_Filtered <- GEO_phenotypes_mGSE69502

saveRDS(GEO_phenotypes_Filtered,
     file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_phenotypes410_Filtered.rds")

# saveRDS(GEO_phenotypes_Filtered,
#      file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_phenotypes410_Filtered_add131.rds")

```


### read array datas
```{r, normalPlacenta}
  
# read Red-Green set for 450K samples
GEO_phenotypes_450k <- GEO_phenotypes_Filtered[GEO_phenotypes_Filtered$ArrayType=='450K',]


GEO_phenotypes_450k_normalPlacenta <- GEO_phenotypes_450k %>% 
  filter(str_detect(Sample, pattern="^Placenta$")) %>% 
  filter(str_detect(Disease, pattern="Uncomplicated$"))

# there are `nrow(GEO_phenotypes_450k_normalPlacenta)` normal placenta samples in total
GEO_RGset_450k_normalPlacenta <- read.metharray.exp(base = GEO_baseDir,
                                                    targets = GEO_phenotypes_450k_normalPlacenta,
                                                    extended = TRUE, force = TRUE)
GEO_phenoData_450k_normalPlacenta <- pData(GEO_RGset_450k_normalPlacenta) #chain data & phenotypes together


# read RG set fot EPIC samples
## read our study
GEO_phenotypes_EPIC_ourStudy <- GEO_phenotypes_Filtered[GEO_phenotypes_Filtered$ArrayType%in%'EPIC' & GEO_phenotypes_Filtered$Study%in%'ourStudy',]
## there are 10 normal placenta samples in total random selected from our study 
GEO_RGset_EPIC_normalPlacenta_ourStudy <- read.metharray.exp(base = GEO_baseDir,
                                                             targets = GEO_phenotypes_EPIC_ourStudy,
                                                             extended = TRUE, force = TRUE)

GEO_phenoData_EPIC_normalPlacenta_ourStudy <- pData(GEO_RGset_EPIC_normalPlacenta_ourStudy)

## read other EPIC data, except our study
GEO_phenotypes_EPIC <- GEO_phenotypes_Filtered[GEO_phenotypes_Filtered$ArrayType%in%'EPIC' & !(GEO_phenotypes_Filtered$Study%in%'ourStudy'),]

GEO_phenotypes_EPIC_normalPlacenta <- GEO_phenotypes_EPIC %>% 
  filter(str_detect(Sample, pattern="^Placenta$")) %>% 
  filter(str_detect(Disease, pattern="Uncomplicated$"))

## there are 25 normal placenta samples in total, except our 10 samples
GEO_RGset_EPIC_normalPlacenta <- read.metharray.exp(base = GEO_baseDir,
                                                    targets = GEO_phenotypes_EPIC_normalPlacenta,
                                                    extended = TRUE, force = TRUE)
GEO_phenoData_EPIC_normalPlacenta <- pData(GEO_RGset_EPIC_normalPlacenta) #chain data & phenotypes together

# save all outputs
save(GEO_RGset_450k_normalPlacenta, GEO_phenoData_450k_normalPlacenta, 
     GEO_RGset_EPIC_normalPlacenta, GEO_phenoData_EPIC_normalPlacenta, 
     GEO_RGset_EPIC_normalPlacenta_ourStudy, GEO_phenoData_EPIC_normalPlacenta_ourStudy,
     file ="/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_RG&pData_normalPlacenta.RData")

```

Check whether we can read 450k and EPIC data together: the result shows that we can't read in the 450k and EPIC array together. So we probably need to account for the batch effect between these 2 platforms.

<!-- ```{r} -->

<!-- GEO_phenotypes_normalPlacenta <- GEO_phenotypes_Filtered %>% -->
<!--    filter(str_detect(Sample, pattern="^Placenta$")) %>% -->
<!--    filter(str_detect(Disease, pattern="Uncomplicated$")) -->

<!-- GEO_RGset_normalPlacenta <- read.metharray.exp(base = GEO_baseDir, targets = GEO_phenotypes_normalPlacenta, -->
<!--                                                force = TRUE) -->

<!-- ``` -->


# read other tissue samples, also uncomplicated samples  

 * GSE74738, 450K data, could be gold standard for different tissue types.

### functions for getting RGset--the raw data set for each tissue types.

we write 2 functions, one for getting RGset form 450k data, another for getting RGset from EPIC data.

```{r function2ReadFiles}

GEOmeta$Sample %>% table()
GEO_phenotypes_450k$Sample %>% table()
GEO_phenotypes_EPIC$Sample %>% table()
GEO_phenotypes_EPIC_ourStudy$Sample %>% table()

# 2 functions to get RGset from 450K and EPIC platform:
## get normal tissue data, 450K
getRGset450K <- function(Tissue){
  # base directory
  GEO_baseDir <- file.path("/home/qianhui/DNAme/Process_decidual/GEO_dataSets/IDAT_all")
  # get 450k meta data
  GEO_phenotypes_450k <- GEO_phenotypes_Filtered[GEO_phenotypes_Filtered$ArrayType=='450K',]
  # get 450k, meta data for the tissue
  GEO_phenotypes_450k_Tissue <- GEO_phenotypes_450k %>% 
  filter(str_detect(Sample, pattern=paste0("^", Tissue, "$"))) %>% 
  filter(str_detect(Disease, pattern="Uncomplicated$"))
  # get the RGset for this tissue
  GEO_RGset_450k_Tissue <- read.metharray.exp(base = GEO_baseDir,
                                              targets = GEO_phenotypes_450k_Tissue,
                                              extended = TRUE, force = TRUE)
  # GEO_phenoData_450k_Tissue <- pData(GEO_RGset_450k_Tissue) #chain data & phenotypes together
  
  return(GEO_RGset_450k_Tissue)
}

## get normal tissue data, EPIC
getRGsetEPIC <- function(Tissue){
  # base directory
  GEO_baseDir <- file.path("/home/qianhui/DNAme/Process_decidual/GEO_dataSets/IDAT_all")
  # get 450k meta data
  GEO_phenotypes_EPIC <- GEO_phenotypes_Filtered[GEO_phenotypes_Filtered$ArrayType=='EPIC',]
  # get 450k, meta data for the tissue
  GEO_phenotypes_EPIC_Tissue <- GEO_phenotypes_EPIC %>% 
  filter(str_detect(Sample, pattern=paste0("^", Tissue, "$"))) %>% 
  filter(str_detect(Disease, pattern="Uncomplicated$"))
  # get the RGset for this tissue
  GEO_RGset_EPIC_Tissue <- read.metharray.exp(base = GEO_baseDir,
                                              targets = GEO_phenotypes_EPIC_Tissue,
                                              extended = TRUE, force = TRUE)
  # GEO_phenoData_450k_Tissue <- pData(GEO_RGset_450k_Tissue) #chain data & phenotypes together
  
  return(GEO_RGset_EPIC_Tissue)
}


```


### other 5tissue types:
The other 5tissue types except for placenta, including:
Amnoin, Chorion, Decidua, maternal whole blood, placental mesenchyme, placental trophoblast, umbilical cord and umbilical cord blood.

#### 1) Amnion
```{r Amnion}
# Amnion
## 450k
GEO_RGset_450k_Amnion <- getRGset450K(Tissue = "Amnion")
GEO_phenoData_450k_Amnion <- pData(GEO_RGset_450k_Amnion) #chain data & phenotypes together
## EPIC
# GEO_RGset_EPIC_Amnion <- getRGsetEPIC(Tissue = "Amnion")
# GEO_phenoData_EPIC_Amnion <- pData(GEO_RGset_EPIC_Amnion) #chain data & phenotypes together
## save outputs for this tissue
save(GEO_RGset_450k_Amnion, GEO_phenoData_450k_Amnion, 
     # GEO_RGset_EPIC_Amnion, GEO_phenoData_EPIC_Amnion, 
     file ="/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_RG&pData_normalAmnion.RData")

```

#### 2) Chorion
```{r Chorion}
## 450k
GEO_RGset_450k_Chorion <- getRGset450K(Tissue = "Chorion")
GEO_phenoData_450k_Chorion <- pData(GEO_RGset_450k_Chorion) #chain data & phenotypes together
## EPIC
# GEO_RGset_EPIC_Chorion <- getRGsetEPIC(Tissue = "Chorion")
# GEO_phenoData_EPIC_Chorion <- pData(GEO_RGset_EPIC_Chorion) #chain data & phenotypes together
## save outputs for this tissue
save(GEO_RGset_450k_Chorion, GEO_phenoData_450k_Chorion, 
     # GEO_RGset_EPIC_Chorion, GEO_phenoData_EPIC_Chorion, 
     file ="/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_RG&pData_normalChorion.RData")

```


#### 3) Decidua
```{r Decidua}
## 450k
GEO_RGset_450k_Decidua <- getRGset450K(Tissue = "Decidua")
GEO_phenoData_450k_Decidua <- pData(GEO_RGset_450k_Decidua) #chain data & phenotypes together

## EPIC
GEO_RGset_EPIC_Decidua <- getRGsetEPIC(Tissue = "Decidua")
GEO_phenoData_EPIC_Decidua <- pData(GEO_RGset_EPIC_Decidua) #chain data & phenotypes together
## save outputs for this tissue
save(GEO_RGset_450k_Decidua, GEO_phenoData_450k_Decidua, 
     GEO_RGset_EPIC_Decidua, GEO_phenoData_EPIC_Decidua, 
     file ="/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_RG&pData_normalDecidua.RData")

```

#### 4) Maternal whole blood
```{r MaternalBlood}
## 450k
GEO_RGset_450k_MaternalBlood <- getRGset450K(Tissue = "Maternal whole blood")
GEO_phenoData_450k_MaternalBlood <- pData(GEO_RGset_450k_MaternalBlood) #chain data & phenotypes together

## save outputs for this tissue
save(GEO_RGset_450k_MaternalBlood, GEO_phenoData_450k_MaternalBlood, 
     # GEO_RGset_EPIC_MaternalBlood, GEO_phenoData_EPIC_MaternalBlood, 
     file ="/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_RG&pData_normalMaternalBlood.RData")

```


#### 5) Umbilical cord blood

```{r CordBlood}

## 450k
GEO_RGset_450k_UCB <- getRGset450K(Tissue = "Umbilical cord blood")
GEO_phenoData_450k_UCB <- pData(GEO_RGset_450k_UCB) #chain data & phenotypes together

## save outputs for this tissue
save(GEO_RGset_450k_UCB, GEO_phenoData_450k_UCB, 
     file ="/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_RG&pData_normalUmbilicalCordBlood.RData")

```

