---
title: "technical report (mainly 450k data, 15 GEO data sets used)"
author: "Qianhui"
date: "21/11/2018"
output: html_document
---

This document is mainly for part3 of the GEO data set analysis, that is the quality control of data sets from different tissue types.

I did PCA on control porbes, so we can see whether there are bach effects between different studies, but within same type of tissue.


```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", error = TRUE, echo = TRUE) # fig.width = 7, 
```


# 1. load data and packages
```{r package, warning=FALSE, message=FALSE}

library(tidyverse)
library(minfi)
library(magrittr)
library(tibble)
library(data.table)
library(limma)
library(ewastools)
library(FactoMineR)
library(doParallel)
library(ggplot2)
library(RColorBrewer)
pal <- brewer.pal(8, "Dark2")

# GEOmeta, GEO_phenotypes
load(file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_phenotypes_add10.RData")

#GEO_RGset_450k_normalPlacenta, GEO_phenoData_450k_normalPlacenta, 
#GEO_RGset_EPIC_normalPlacenta, GEO_phenoData_EPIC_normalPlacenta, 
#GEO_RGset_EPIC_normalPlacenta_ourStudy, GEO_phenoData_EPIC_normalPlacenta_ourStudy
load(file ="/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_RG&pData_normalPlacenta.RData")

#GEO_RGset_450k_Amnion, GEO_phenoData_450k_Amnion, GEO_RGset_EPIC_Amnion, GEO_phenoData_EPIC_Amnion
load(file ="/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_RG&pData_normalAmnion.RData")

#GEO_RGset_450k_Chorion, GEO_phenoData_450k_Chorion,GEO_RGset_EPIC_Chorion, GEO_phenoData_EPIC_Chorion
load(file ="/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_RG&pData_normalChorion.RData")

#GEO_RGset_450k_Decidua, GEO_phenoData_450k_Decidua, GEO_RGset_EPIC_Decidua, GEO_phenoData_EPIC_Decidua, 
load(file ="/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_RG&pData_normalDecidua.RData")

#GEO_RGset_450k_MaternalBlood, GEO_phenoData_450k_MaternalBlood, 
load(file ="/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_RG&pData_normalMaternalBlood.RData")

#GEO_RGset_450k_mesenchyme, GEO_phenoData_450k_mesenchyme
load(file ="/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_RG&pData_normalPlacentalMesenchyme.RData")

#GEO_RGset_450k_trophoblast, GEO_phenoData_450k_trophoblast
load(file ="/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_RG&pData_normalPlacentalTrophoblast.RData")

#GEO_RGset_450k_UC, GEO_phenoData_450k_UC
load(file ="/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_RG&pData_normalUmbilicalCord.RData")

#GEO_RGset_450k_UCB, GEO_phenoData_450k_UCB
load(file ="/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_RG&pData_normalUmbilicalCordBlood.RData")

```


# 2. look at PCA of control probes

## Function for control probe PCA, to identify potential batch effects.
```{r PCAofControlProbes}
# let argument RGset to be a charactor
# argument `group`, group could be `trimester`, `fetalSex` `gestationalAge` or `Study`  

ControlPCA <- function(RGsetChara, group){
  registerDoParallel(cores = 10)
  
  # get the object from character
  RGset <- get(RGsetChara)
  # get meta data
  phenoData <- pData(RGset)

  # get intensity of control probes:
  ctrls <- getProbeInfo(RGset, type = "Control")

  M.Neg <- getGreen(RGset)[ctrls$Address[ctrls$Type == "NEGATIVE"], , drop=FALSE]
  U.Neg <- getRed(RGset)[ctrls$Address[ctrls$Type == "NEGATIVE"], , drop=FALSE]
  
  # calculate beta values and M values
  beta_ctrls <- M.Neg/(M.Neg+U.Neg+100) %>% as.matrix()
  M_ctrls <- log2(beta_ctrls/(1 - beta_ctrls))
  M_ctrls_v1 <- M_ctrls[apply(M_ctrls, 1, function(x) all(is.finite(x))), , drop=FALSE]

  # remove infinite values and draw PCA plots for data sets with mutiple samples
  if(ncol(M_ctrls)==1){
    
    print("no PCA plot, since only 1 sample for this data set")
    
  }else if(ncol(M_ctrls)==2){
    
    print("no PCA plot, since only 2 samples for this data set")
    
  }else{

  ## PCA plot for control M values
  PCA_M_ctrl <- FactoMineR::PCA(base::t(M_ctrls_v1))

  ## relabel PCA_M_ctrl plot using ggplot:
  PCA_M_ctrl_plot.df <- data.frame(x = PCA_M_ctrl$ind$coord[,1], y = PCA_M_ctrl$ind$coord[,2],
                                 fetalSex=phenoData$Fetal_Sex,
                                 gestationalAge=phenoData$Gestation,
                                 trimester=phenoData$Trimester,
                                 Study= phenoData$Study)
  
  PCA_M_ctrl_plot <-
  PCA_M_ctrl_plot.df %>%
    ggplot(aes(x = x, y = y)) +
  ## add argument `group`, group could be `trimester`, `fetalSex` `gestationalAge` or `Study`  
    geom_point(aes(col = get(group)), alpha = 1, size = 7)+ #color = factor(Batch),
    scale_colour_hue()+
    labs(x="PCA1", y="PCA2", colour=group)+
    ggtitle(RGsetChara)+
    theme(text = element_text(size=20))
  
  print(PCA_M_ctrl_plot)
    
  } # end if
  
  # values to return
  ControlBMList <- list(ControlB=beta_ctrls, ControlM=M_ctrls_v1)
  return(ControlBMList)
  # saveRDS(PCA_M_ctrl, file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/PCA_M_ctrl.rds")
}

```


## Use FUN -- `ControlPCA` for each tissue types: 450K & EPIC (data from GEO data base)
```{r ControlPCAforAllTissues}
# !(GEO_phenotypes$Sample_name%in%c("GSM1617002_7668610068_R01C01","GSM1616993_7668610068_R06C02"))

GEO_RGset_450k_normalPlacenta <- GEO_RGset_450k_normalPlacenta[,-5]
GEO_RGset_450k_MaternalBlood <- GEO_RGset_450k_MaternalBlood[,-8]

TissueRGsetNames <- c(
                      "GEO_RGset_450k_normalPlacenta", "GEO_RGset_EPIC_normalPlacenta",
                      "GEO_RGset_EPIC_normalPlacenta_ourStudy",
                      "GEO_RGset_450k_Amnion", "GEO_RGset_EPIC_Amnion",
                      "GEO_RGset_450k_Chorion", "GEO_RGset_EPIC_Chorion",
                      "GEO_RGset_450k_Decidua", "GEO_RGset_EPIC_Decidua",
                      "GEO_RGset_450k_MaternalBlood", "GEO_RGset_450k_mesenchyme",
                      "GEO_RGset_450k_trophoblast", 
                      "GEO_RGset_450k_UC", "GEO_RGset_450k_UCB"
                      )


for(i in TissueRGsetNames){
    assign(paste0("ControlBM", str_replace_all(i, "GEO_RGset", "")),
           ControlPCA(RGsetChara = i, group = "Study"))
}

ControlBM_450k_normalPlacenta_plotByTrimester <- ControlPCA(RGsetChara = "GEO_RGset_450k_normalPlacenta", group = "trimester")

save(list=c(paste0("ControlBM", str_replace_all(i, "GEO_RGset", ""))),
     file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/ControlBMof9TissueTypes.RData")

```


# 3. Put all control probes together

## Function `JoinAndPlotAllControl`
```{r joinAllandPlotAll}
ControlBMvalInfo <- tibble(
  TissueRGsetNames = c("GEO_RGset_450k_normalPlacenta", "GEO_RGset_EPIC_normalPlacenta",
                       "GEO_RGset_EPIC_normalPlacenta_ourStudy",
                       "GEO_RGset_450k_Amnion", "GEO_RGset_EPIC_Amnion",
                       "GEO_RGset_450k_Chorion", "GEO_RGset_EPIC_Chorion",
                       "GEO_RGset_450k_Decidua", "GEO_RGset_EPIC_Decidua",
                       "GEO_RGset_450k_MaternalBlood", "GEO_RGset_450k_mesenchyme",
                       "GEO_RGset_450k_trophoblast", 
                       "GEO_RGset_450k_UC", "GEO_RGset_450k_UCB"
                      ),
  ControlBMvalNames = paste0("ControlBM", str_replace_all(TissueRGsetNames, "GEO_RGset", ""))
)


# write the function to join all control informations together
# value-- Beta or M
# TissueBMvalInfo -- a tibble containing BM names

JoinAndPlotAllControl <- 
  function(ControlBMvalInfo, value, GEO_phenotypes){
  registerDoParallel(cores = 10)
  # list
  ListToJoin <- list()

  # names for the list
  listElements <- paste0(ControlBMvalInfo$ControlBMvalNames, ".df")
  # fill the list with beta/M values
  if(value=="Beta"){
    for(Name in listElements){
      TissueBMvalNames <- str_replace_all(Name, ".df", "")
      ListToJoin[[Name]] <- get(TissueBMvalNames)$ControlB %>% as.data.frame() %>% rownames_to_column(var="Probe")
    }
    
  }else if(value=="M"){
    for(Name in listElements){
      TissueBMvalNames <- str_replace_all(Name, ".df", "")
      ListToJoin[[Name]] <- get(TissueBMvalNames)$ControlM %>% as.data.frame() %>% rownames_to_column(var="Probe")
    }
  } # end if

  # left_join values form all samples
  Join.df <- ListToJoin %>% purrr::reduce(left_join, by = "Probe")
  # omit na values
  Join.df_v1 <- na.omit(Join.df)
  # as matrix
  Join.df_v2 <- Join.df_v1 %>% remove_rownames() %>% 
  column_to_rownames(var="Probe") %>% as.matrix() # 
  
  #plotMDS plots
  ## phenodata for these samples
  GEO_phenotypes_551 <- GEO_phenotypes[match(colnames(Join.df_v2),
                                                 GEO_phenotypes$Sample_name),]
  ## check names all in same order or not
  table(colnames(Join.df_v2)==GEO_phenotypes_551$Sample_name) %>% print()

  ## plot MDS plot
  par(mfrow=c(1,1))
  MDS_control <- plotMDS(Join.df_v2, top=nrow(Join.df_v2), labels= GEO_phenotypes_551$Sample, gene.selection="common") 

  # plot PCA
   PCA_control <- FactoMineR::PCA(base::t(Join.df_v2))
  
  # return Join.df_v2, GEO_phenotypes_norm552 and MDS_norm552
  output_List <- list(JoinAllControl=Join.df_v2, phenoDataAll=GEO_phenotypes_551, 
                      MDS_control=MDS_control, PCA_control=PCA_control)
  return(output_List)
}

```


## use FUN `JoinAndPlotAllControl`
```{r}

Control551_Beta <- JoinAndPlotAllControl(ControlBMvalInfo =  ControlBMvalInfo, 
                                         value = "Beta", 
                                         GEO_phenotypes = GEO_phenotypes)

Control551_M <- JoinAndPlotAllControl(ControlBMvalInfo =  ControlBMvalInfo, 
                                      value = "M", 
                                      GEO_phenotypes = GEO_phenotypes)

save(Control551_Beta, Control551_M,
     file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/Controls_BM_all551.RData")
```



