---
title: "PCA"
author: "Qianhui"
date: "12/17/2018"
output: html_document
---

This document is mainly for part5 of the GEO data set analysis, that is draw PCA plots

```{r setup, include=FALSE}

knitr::opts_chunk$set(fig.align = "center", error = TRUE, echo = TRUE) # fig.width = 7, 

```

# load package and files
```{r}

library(tidyverse)
library(magrittr)
library(tibble)
library(FactoMineR)
library(doParallel)
library(ggplot2)
library(ggpubr)
library(ggfortify)
library(mclust)
library(cluster)
library(RColorBrewer)
theme_set(theme_pubr())

registerDoParallel(cores=10)

# load data 
## GEOmeta, GEO_phenotypes
load(file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_phenotypes_add10.RData")


PCA_Placenta_M <- readRDS(file  = "/home/qianhui/DNAme/Process_decidual/RDSfiles/PCA_all.Rds")

# NormBMlist <- readRDS(file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/NormBMlist.rds")
NormBMlist <- readRDS(file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/NormBMlist_300k.rds")

## annotation needed for using functions
ann450k <- readRDS(file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/ann450k_GEO15sets.rds")
annEPIC <- readRDS(file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/annEPIC_GEO15sets.rds")

```


# scree plot for PCAs
```{r}

eigenvalues <- PCA_Placenta_M$eig[1:10, ]

tiff(file = paste0("/home/qianhui/DNAme/Process_decidual/figures/ScreePlot.tiff"), 
     width = 20,height = 15, units="cm",  res = 300)

barplot(eigenvalues[, 2], names.arg=1:nrow(eigenvalues), 
       main = "Variances",
       xlab = "Principal Components",
       ylab = "Percentage of variances",
       col ="steelblue")

# Add connected line segments to the plot
lines(x = 1:nrow(eigenvalues), eigenvalues[, 2], 
      type="b", pch=19, col = "red")

dev.off()

```


```{r}

table(colnames(NormBMlist$NormBAll) %in% NormBMlist$phenoDataAll$Sample_name)
Matched_pd <- NormBMlist$phenoDataAll
Matched_pd$Outlier <- ifelse(Matched_pd$Additional_Info%in%"Outlier"|Matched_pd$Sample_name%in%c("GSM1702177_7970368142_R05C02", "GSM1947213_6008581028_R01C01"),
                             "Outlier", "Standard")

NormMAll <- NormBMlist$NormMAll

# T1 PCA
PhenodataPlacentaT1 <- Matched_pd[Matched_pd$Sample%in%"Placenta" & Matched_pd$Trimester%in%"First",] 
Placenta_Mnorm_T1 <-  NormMAll[,colnames(NormMAll)%in%PhenodataPlacentaT1$Sample_name]

table(PhenodataPlacentaT1$Sample_name==colnames(Placenta_Mnorm_T1))
# PCA_PlacentaT1_M <- FactoMineR::PCA(base::t(Placenta_Mnorm_T1))

# T2 PCA
PhenodataPlacentaT2 <- Matched_pd[Matched_pd$Sample%in%"Placenta" & Matched_pd$Trimester%in%"Second",] 
Placenta_Mnorm_T2 <-  NormMAll[,colnames(NormMAll)%in%PhenodataPlacentaT2$Sample_name]

table(PhenodataPlacentaT2$Sample_name==colnames(Placenta_Mnorm_T2))
# PCA_PlacentaT2_M <- FactoMineR::PCA(base::t(Placenta_Mnorm_T2))

# Term PCA
PhenodataPlacentaTerm <- Matched_pd[Matched_pd$Sample%in%"Placenta" &Matched_pd$Trimester%in%"Term",] 
Placenta_Mnorm_Term <-  NormMAll[,colnames(NormMAll)%in%PhenodataPlacentaTerm$Sample_name]

table(PhenodataPlacentaTerm$Sample_name==colnames(Placenta_Mnorm_Term))
# PCA_PlacentaTerm_M <- FactoMineR::PCA(base::t(Placenta_Mnorm_Term))

# save(PCA_Placenta_M, PCA_PlacentaT1_M, PCA_PlacentaT2_M, PCA_PlacentaTerm_M, file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/PCA_placenta_deciduals_allT1Term_BMval.RData")
# load(file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/PCA_placenta_deciduals_allT1Term.RData")

```


# `ggplotPCA` function:
```{r}

# write the ggplot PCA function:================================
ggplotPCA <- function(PCA_Placenta_M, Matched_pd, value){
  
new_plotDF <- tibble(x=PCA_Placenta_M$ind$coord[,1], 
                     y=PCA_Placenta_M$ind$coord[,2],
                     GA = Matched_pd$Gestation,
                     FetalSex= Matched_pd$Fetal_Sex,
                     SampleNames= Matched_pd$Sample_name,
                     Tissue0= Matched_pd$Sample,
                     Tissue=str_replace_all(Matched_pd$Sample, c(
                            ".lacenta.+"="Placenta",
                            "umbilical"="Umbilical")),
                     Trimester=Matched_pd$Trimester,
                     Study=Matched_pd$Study,
                     Outlier=ifelse(Matched_pd$Additional_Info%in%"Outlier", "Outlier", "Pure"),
                     ArrayType=Matched_pd$ArrayType) %>% 
  mutate(Outlier2 = case_when(SampleNames %in% c("GSM1702177_7970368142_R05C02", "GSM1947213_6008581028_R01C01") ~ "Outlier",
                             !(SampleNames %in% c("GSM1702177_7970368142_R05C02", "GSM1947213_6008581028_R01C01")) ~ Outlier))
  
colnames(new_plotDF) <- c("x", "y", "Gestational age", "FetalSex", "SampleNames",
                          "Tissue0","Tissue", 
                          "Trimester", "Study", "Outlier", "ArrayType", "Outlier2")

# new_plotDF$Tissue[which(new_plotDF$Outlier2=="Outlier")] <- rep("Placental outliers", 11)
new_plotDF$Tissue[which(new_plotDF$Outlier2=="Outlier")] <- rep("Impure placenta", 11)

# Plot the new dataframe
## color to use

PCAplot_norm408 <- new_plotDF %>%
  ggplot(aes(x, y)) + 
  geom_point(aes(shape = Tissue, col=Trimester), alpha = 1, size = 5)+
  # geom_point(aes(shape = Trimester, col=Tissue), alpha = 1, size = 5)+ 
  # scale_color_manual(values = namedPal2)+
  scale_shape_manual(values=1:11)+
  # scale_color_manual(name="Tissue", values = c("#000000", "#BD0026", "#984EA3", "#999999", "#D95F02", "#1B9E77", "#F781BF", "#A65628"))+
  labs(x="PC1", y="PC2")+
  theme(text = element_text(size=20))
# axis.text.x = element_text(angle=90, hjust=1)) 

# save the MDS plot as pdf:
pdf(file = paste0("/home/qianhui/DNAme/Process_decidual/figures/MDSplot-Normalised", value, "_joinedGEOsamples408.pdf"),
    width = 13, height = 7)

# theme_set(theme_bw())
PCAplot_norm408
dev.off()

# show the plot
# theme_set(theme_bw())
PCAplot_norm408 %>% print()

# return new_plotDF
return(new_plotDF)
}
# end of ggplotPCA function=======================================

```



# use ggplotPCA function to plot PCAs
```{r}

## all samples
PCA408 <- ggplotPCA(PCA_Placenta_M = PCA_Placenta_M, Matched_pd = Matched_pd, value = "all408NormMAll")

PCA408$TissueNoGreen <- str_replace_all(PCA408$Tissue, c("Impure placenta"="Placenta"))

PCA408plot <- PCA408 %>%
  ggplot(aes(x, y)) + 
  # geom_point(aes(shape = Tissue, col=Trimester), alpha = 1, size = 5)+ 
  geom_point(aes(shape = Trimester, col=TissueNoGreen), alpha = 1, size = 2)+ 
  scale_color_manual(name="Tissue", values = c("Amnion"="#000000", "Chorion"="#BD0026", 
                                               "Decidua"="#984EA3", "Maternal whole blood"="#999999", 
                                               "Placenta"="#D95F02", #"Impure placenta"="#1B9E77", 
                                               "Umbilical cord blood"="#F781BF"))+
  scale_shape_manual(values=c(0,1,2))+
  labs(x="PC1 (25.91%)", y="PC2 (10.69%)")+
  scale_y_continuous(breaks=c(-600,-300,0, 300, 600), limits = c(-600, 800))+
  scale_x_continuous(breaks=c(-400,0, 400, 800), limits = c(-410, 1150))+
  # ylim(-600,600)+
  # xlim(-410, 1150)+
  guides(color=guide_legend(nrow = 3), shape=guide_legend(nrow = 2)) +
  # stat_ellipse(aes(x, y, col=Tissue0), type = "t") +
  # stat_ellipse(type = "t")+ #type = "norm", linetype = 2
  geom_vline(aes(xintercept=quantile(PCA_Placenta_M$ind$coord[, 1], 0.75)+ 1.5*IQR(PCA_Placenta_M$ind$coord[, 1])),
             linetype=4, colour="black")+
  geom_vline(aes(xintercept=quantile(PCA_Placenta_M$ind$coord[, 1], 0.75)+ 3*IQR(PCA_Placenta_M$ind$coord[, 1])),
             linetype=4, colour="black")+
  # geom_vline(aes(xintercept=quantile(PCA_Placenta_M$ind$coord[, 1], 0.25)- 1.5*IQR(PCA_Placenta_M$ind$coord[, 1])),
  #            linetype=4, colour="black")+
  # geom_text(aes(x=155 , y=-600, label="Q3+1.5*IQR"), vjust=-0.4, hjust=0, size=3) + #angle=90,, "Q3+3*IQR")
  # geom_text(aes(x=300 , y=-600, label="Q3+3*IQR"), vjust=-0.4, hjust=0, size=3) + #angle=90,, "Q3+3*IQR")
  theme(text = element_text(size=9),
        legend.title=element_text(size=9),
        legend.text=element_text(size=9),
        axis.text.x =element_text(size=9, colour = "black"),
        axis.text.y =element_text(size=9, colour = "black"),
        legend.position="top")

tiff(file = paste0("/home/qianhui/DNAme/Process_decidual/figures/MDSplot-Normalised", "allPCA408", "_joinedGEOsamples408.tiff"), 
     width = 20,height = 15, units="cm",  res = 300)
# theme_set(theme_bw())
PCA408plot
dev.off()



# ## T1 sampels
# PCAT1 <- ggplotPCA(PCA_Placenta_M = PCA_PlacentaT1_M, Matched_pd = PhenodataPlacentaT1, value = "all_T1_NormMAll")
# ## Term sampels
# PCATerm <- ggplotPCA(PCA_Placenta_M = PCA_PlacentaTerm_M, Matched_pd = PhenodataPlacentaTerm, value = "all_Term_NormMAll")

# rbind df and plot them together by facet_wrap

All408rbind <- PCA408
All408rbind[is.na(All408rbind$FetalSex), ]$FetalSex <- rep("Female",2)

All408rbind$TrimesterSex <- do.call(paste0, All408rbind[c("Trimester", "FetalSex")])

PCAallplot <- All408rbind %>%
  ggplot(aes(x, y)) + 
  geom_point(aes(shape = Trimester, col=TissueNoGreen), alpha = 1, size = 2)+ 
  scale_color_manual(name="Tissue", values = c("Amnion"="#000000", "Chorion"="#BD0026", 
                                               "Decidua"="#984EA3",  
                                               "Placenta"="#D95F02", "Impure placenta"="#1B9E77", 
                                               "Umbilical cord blood"="#F781BF","Maternal whole blood"="#999999"))+
  scale_shape_manual(values=c(0,1,2))+
  # scale_shape_manual(values=c(FirstFemale=0,FirstMale=15, 
  #                             SecondFemale=1,SecondMale=16, 
  #                             TermFemale=2, TermMale=17))+
  labs(x="PC1", y="PC2")+
  scale_y_continuous(breaks=c(-600,-300,0, 300, 600), limits = c(-600, 800))+
  scale_x_continuous(breaks=c(-400,0, 400, 800), limits = c(-410, 1150))+
  guides(color=guide_legend(nrow = 3), shape=guide_legend(nrow = 2)) +
  # stat_ellipse(data =All408rbind, aes(x, y,col=Tissue),
  #              type = "norm", show.legend =FALSE, inherit.aes = FALSE, level = 0.8) +
  geom_vline(aes(xintercept=quantile(PCA_Placenta_M$ind$coord[, 1], 0.75)+ 1.5*IQR(PCA_Placenta_M$ind$coord[, 1])),
             linetype=4, colour="black")+
  geom_vline(aes(xintercept=quantile(PCA_Placenta_M$ind$coord[, 1], 0.75)+ 3*IQR(PCA_Placenta_M$ind$coord[, 1])),
             linetype=4, colour="black")+
  facet_wrap(~Trimester)+
  theme(text = element_text(size=9),
        axis.text.x =element_text(size=9, colour = "black"),
        legend.title=element_text(size=6),
        legend.text=element_text(size=6),
        legend.position="top")
```


plot study batches:
```{r}

colorFun <- colorRampPalette(brewer.pal(9, "Set1"))
pal2 <- colorFun(15)
namedPal2 <- pal2 %>% `names<-`(unique(PCA408$Study))
# namedPal2 <- pal2 %>% `names<-`(unique(PCA408$Trimester))
# namedPal2 <- pal2 %>% `names<-`(unique(PCA408$FetalSex))



PCAplot408 <- PCA408 %>%
  ggplot(aes(x, y)) + 
  geom_point(aes(col = Study), alpha = 1, size = 5)+
  # geom_point(aes(col=Trimester), alpha = 1, size = 5)+
  # geom_point(aes(col=FetalSex), alpha = 1, size = 5)+
  # scale_colour_gradient(low = "#56B1F7", high = "#132B43")+
  # scale_color_brewer(palette = "Set1")
  scale_color_manual(values = namedPal2)+
  # scale_shape_manual(values=1:length(new_plotDF$Tissue))+
  labs(x="PC1 (25.91%)", y="PC2 (10.69%)")+
  theme(text = element_text(size=15),
        legend.position="right")


tiff(file = paste0("/home/qianhui/DNAme/Process_decidual/figures/ColorByStudyBatches.tiff"), 
     width = 20,height = 15, units="cm",  res = 300)
theme_set(theme_pubr())
PCAplot408
dev.off()
```

save files:
```{r}
saveRDS(PCA408, file = "/home/qianhui/DNAme/Process_decidual/figures/PCA408df.rds")

saveRDS(PCA408plot, file = "/home/qianhui/DNAme/Process_decidual/figures/T1TermSamples_PCA408samples.rds")

# saveRDS(PCAallplot, file = "/home/qianhui/DNAme/Process_decidual/figures/PCAallplot_noSexLabel.rds")

saveRDS(PCAallplot, file = "/home/qianhui/DNAme/Process_decidual/figures/PCAallplot.rds")

```


Compare with results from ewastool R package:
```{r}


GEO_phenotypes_ewastoolPurityCheck_output <- read_csv(file =  "/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_phenotypes_ewastoolPurityCheck_output.csv")

ewastoolOutlier <- GEO_phenotypes_ewastoolPurityCheck_output[GEO_phenotypes_ewastoolPurityCheck_output$log_odds> -4,]$Sample_name

PCA408$ewastoolOutlier <- ifelse(PCA408$SampleNames%in%ewastoolOutlier, "ewastoolOutlier", "ewastoolStandard")


PCA408 %>%
  ggplot(aes(x, y)) + 
  geom_point(aes(shape = TissueNoGreen, col=ewastoolOutlier), alpha = 1, size = 2)+ 
  # scale_color_manual(name="Tissue", values = c("Amnion"="#000000", "Chorion"="#BD0026",
  #                                              "Decidua"="#984EA3", "Maternal whole blood"="#999999", 
  #                                              "Placenta"="#D95F02", #"Impure placenta"="#1B9E77", 
  #                                              "Umbilical cord blood"="#F781BF"))+
  scale_shape_manual(values=c("Amnion"=0, "Chorion"=1, "Decidua"=2, "Maternal whole blood"=3, 
                                   "Placenta"=4, "Umbilical cord blood"=5))+
  labs(x="PC1 (25.91%)", y="PC2 (10.69%)", shape="Tissue")+
  scale_y_continuous(breaks=c(-600,-300,0, 300, 600), limits = c(-600, 800))+
  scale_x_continuous(breaks=c(-400,0, 400, 800), limits = c(-410, 1150))+
  guides(color=guide_legend(nrow = 3), shape=guide_legend(nrow = 2)) +
  geom_vline(aes(xintercept=quantile(PCA_Placenta_M$ind$coord[, 1], 0.75)+ 1.5*IQR(PCA_Placenta_M$ind$coord[, 1])),
             linetype=4, colour="black")+
  geom_vline(aes(xintercept=quantile(PCA_Placenta_M$ind$coord[, 1], 0.75)+ 3*IQR(PCA_Placenta_M$ind$coord[, 1])),
             linetype=4, colour="black")+
  theme(text = element_text(size=9),
        legend.title=element_text(size=9),
        legend.text=element_text(size=9),
        axis.text.x =element_text(size=9, colour = "black"),
        axis.text.y =element_text(size=9, colour = "black"),
        legend.position="top")

```


# mclust for classifying samples based on PCA plot
```{r}
# all saples PCA
NormMAll <- NormBMlist$NormMAll
table(colnames(NormMAll) %in% GEO_phenotypes$Sample_name)

Matched_pd <- GEO_phenotypes[match(colnames(NormMAll), GEO_phenotypes$Sample_name),]

df <- PCA_Placenta_M$ind$coord %>% as.data.frame() %>% rownames_to_column(var="Sample_name") %>% 
  left_join(Matched_pd, by="Sample_name") %>% column_to_rownames(var="Sample_name")

a1 <- mclust::Mclust(df[,c("Dim.1","Dim.2")],2) #for 2 groups
table(a1$classification,df$Sample)
with(df,plot(Dim.2 ~ Dim.1,pch=as.numeric(as.factor(Sample)),col=a1$classification))
with(df,legend("topright",levels(as.factor(Sample)),pch=1:6))

# plot mclust result
df$class <- a1$classification

PCAclassPlot <- df %>%
  ggplot(aes(Dim.1, Dim.2)) + 
  geom_point(aes(shape = Sample, col=as.factor(class)), alpha = 1, size = 2)+ 
  scale_shape_manual(values=c(0:5))+
  scale_color_manual(name="Clusters", values = c("1"="#000000","2"="#D95F02"), 
                     labels=c("1"="Cluster 1", "2"="Cluster 2"))+
  labs(x="PC1 (25.91%)", y="PC2 (10.69%)")+
  scale_y_continuous(breaks=c(-600,-300,0, 300, 600), limits = c(-600, 800))+
  scale_x_continuous(breaks=c(-400,0, 400, 800), limits = c(-410, 1150))+
  # ylim(-600,600)+
  # xlim(-410, 1150)+
  guides(color=guide_legend(nrow = 3), shape=guide_legend(nrow = 2)) +
  theme(text = element_text(size=9),
        legend.title=element_text(size=9),
        legend.text=element_text(size=9),
        axis.text.x =element_text(size=9, colour = "black"),
        axis.text.y =element_text(size=9, colour = "black"),
        legend.position="top")

tiff(file = paste0("/home/qianhui/DNAme/Process_decidual/figures/PCAmclustClassPlot_v1.tiff"), 
     width = 20,height = 15, units="cm",  res = 300)
theme_set(theme_pubr())
PCAclassPlot
dev.off()


```

# mclust output file
Add the detailed outlier information to each outlier:
```{r}

PercClassOutPut <- readRDS(file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/PCALDAresult_PercClassOutPut.rds")

OutlierDetail <- PercClassOutPut %>% 
  dplyr::select(SampleNames, OutlierDetail=Outlier) %>% 
  dplyr::left_join(PCA408[, c("SampleNames", "Outlier2")], by = "SampleNames") %>% 
  dplyr::select(SampleNames, OutlierDetail, Outlier = Outlier2)
  
saveRDS(OutlierDetail, file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/OutlierDetails.rds")

```


```{r}

# output file for the placenta percentage

PercClassOutPutMclust <- a1$z %>% as.data.frame() %>% rownames_to_column(var="Sample_name") %>% 
  left_join(OutlierDetail, by=c("Sample_name"="SampleNames")) %>% 
  left_join(PCA408, by=c("Sample_name"="SampleNames")) %>% 
  dplyr::select(Sample_name, TissueTypes=Tissue, PC1=x, PC2=y, 
                Trimester, Sex=FetalSex, Study, ArrayType, 
                Outlier=Outlier.x,  PlacentaPercentageMclust=V2) %>%
  arrange(PlacentaPercentageMclust)
  

write_csv(PercClassOutPutMclust, path = "/home/qianhui/DNAme/Process_decidual/RDSfiles/PercClassOutPutMclust.csv")

```


Density plot of probabilities of placenta sample being a mixed sample:
```{r}

PercClassOutPutMclust %>% 
  ggplot(aes(x=PlacentaPercentageMclust))+
  geom_density()
  # geom_histogram(binwidth = 0.1, fill = "grey", color = "black")+
  labs(x="Estimated percentage of placenta tissue")

```


Where the potential mixed placenta sample locate:
```{r}

PercClassOutPutMclust$Sample_name[41:44]

df$potencial <- rownames(df)%in%PercClassOutPutMclust$Sample_name[41:44]

df %>%
  ggplot(aes(Dim.1, Dim.2)) + 
  geom_point(aes(shape = as.factor(potencial), col=as.factor(class)), alpha = 1, size = 2)+ 
  scale_shape_manual(values=c(0:5))+
  scale_color_manual(name="Clusters", values = c("1"="#000000","2"="#D95F02"), 
                     labels=c("1"="Cluster 1", "2"="Cluster 2"))+
  labs(x="PC1 (26.11%)", y="PC2 (10.35%)")+
  scale_y_continuous(breaks=c(-600,-300,0, 300, 600), limits = c(-600, 800))+
  scale_x_continuous(breaks=c(-400,0, 400, 800), limits = c(-410, 1150))
  # ylim(-600,600)+
  # xlim(-410, 1150)+

```


# traning data and test data

get M values for all placenta samples:
```{r}

# get placenta data and sample them
table(colnames(NormBMlist$NormMAll)==NormBMlist$phenoDataAll$Sample_name)
datMeth_placenta <- NormBMlist$NormMAll[,NormBMlist$phenoDataAll$Sample%in%"Placenta"]

datMeth_placenta <- NormBMlist$NormMAll

```

Function for dividing, traning and test data:
```{r}

MclustTrainTestFUN <- function(datMeth_placenta){
library(gridExtra)
# Deviding data:
ind <- sample.int(n=ncol(datMeth_placenta), size = floor(0.7*ncol(datMeth_placenta)), replace = FALSE) # floor() is used to return an integer instead of a decimal.

datMethTraining <- datMeth_placenta[,ind]

datMethTesting <- datMeth_placenta[,-ind]

# # plot numbers of outliers in each data set:
# TrainingData=table(colnames(datMethTraining)%in%OutlierDetail[OutlierDetail$Outlier%in%"Outlier",]$SampleNames) %>% 
#     as.data.frame() %>% dplyr::select(Freq)
# 
# TestData=table(colnames(datMethTesting)%in%OutlierDetail[OutlierDetail$Outlier%in%"Outlier",]$SampleNames) %>% 
#   as.data.frame() %>% dplyr::select(Freq) 
# 
# tibble(Samples=c("Pure placenta", "Outlier"), TrainingData,TestData) %>% 
#   grid.table(rows=NULL)

# Training data:
PCA_traning <- PCA(t(datMethTraining))

df_training <- PCA_traning$ind$coord %>% as.data.frame() %>% rownames_to_column(var="Sample_name") %>% 
  left_join(NormBMlist$phenoDataAll, by="Sample_name") %>% column_to_rownames(var="Sample_name")

mclust_training <- mclust::Mclust(df_training[,c("Dim.1","Dim.2")],2) #for 2 groups

table(mclust_training$classification,df_training$Sample)
with(df_training,plot(Dim.2 ~ Dim.1,pch=as.numeric(as.factor(Sample)),col=mclust_training$classification))
with(df_training,legend("topright",levels(as.factor(Sample)),pch=1:6))

df_training$class <- mclust_training$classification

# test data

PCA_test <- predict.PCA(object = PCA_traning, newdata = t(datMethTesting))

df_test <- PCA_test$coord %>% as.data.frame() %>% rownames_to_column(var="Sample_name") %>% 
  left_join(NormBMlist$phenoDataAll, by="Sample_name") %>% column_to_rownames(var="Sample_name")

mclust_test <- predict.Mclust(object = mclust_training, newdata = df_test[,c("Dim.1","Dim.2")])

df_test$class <- mclust_test$classification

return(list(PCA_traning=PCA_traning, mclust_training=mclust_training, mclust_test=mclust_test,
            df_training=df_training, df_test=df_test))

}

```


set seeds and use function
```{r}

# set.seed(2001)
# set.seed(2020)
# set.seed(2021)
set.seed(2022)

res2022 <- MclustTrainTestFUN(datMeth_placenta = datMeth_placenta)

# plot mclust result:
source(file = "~/DNAme/Process_decidual/GEO_TechPaper/Functions/GEOdataAnalysis_TraningTestFuns.R")

ggarrange(
PlotClassFUN(df = res2022$df_training, PCA_traning = res2022$PCA_traning),
PlotClassFUN(df = res2022$df_test, PCA_traning = NULL), common.legend = TRUE)


# TableTrainTestFUN(df=res2022$df_training)
# TableTrainTestFUN(df=res2022$df_test)

```




