---
title: "technical report_use a subset of data"
author: "Qianhui"
date: "21/11/2018"
output: html_document
---

This document is mainly for part4 of the GEO data set analysis, including filtering, dye bias correction, backgroud correction and normalisation steps.

The preprocessed beta values and the correspounding M values were stored in a list. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", error = TRUE, echo = TRUE) # fig.width = 7, 
```


# 1. load data and packages
```{r package, warning=FALSE, message=FALSE}

library(tidyverse)
library(readxl)
library(minfi)
library(magrittr)
library(tibble)
library(data.table)
library(limma)
library(ewastools)
library(ENmix)
library(wateRmelon)
library(FactoMineR)
library(ChAMP)
library(mclust)
library(doParallel)
library(ggplot2)
library(RColorBrewer)
pal <- brewer.pal(8, "Dark2")

# GEOmeta, GEO_phenotypes
load(file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_phenotypes_add10.RData")
GEO_phenotypes_Filtered <- readRDS(file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/GEO_phenotypes410_Filtered.rds")
GEO_phenotypes_Filtered <- GEO_phenotypes_Filtered[!(GEO_phenotypes_Filtered$Sample_name%in%c("GSM1617002_7668610068_R01C01","GSM1616993_7668610068_R06C02")),]


GEO_phenotypes_Filtered_sub <- GEO_phenotypes_Filtered[!(GEO_phenotypes_Filtered$Study%in%
                                                         c("GSE75196", "GSE75248", "GSE100197", 
                                                           "GSE98224", "GSE71678", "GSE115508")),]

# crossreactive probes-- EPIC
crossreactiveProbesEPIC <- read_csv("/home/qianhui/DNAme/Process-QH/13059_2016_1066_MOESM1_ESM-cross-reactive.csv")
colnames(crossreactiveProbesEPIC)[1] <- "ProbeNames"

# crossreactive probes--450k
crossreactiveProbes450k <- read_excel("/home/qianhui/DNAme/Process-QH/48639-non-specific-probes-Illumina450k.xlsx", sheet = 1)
colnames(crossreactiveProbes450k)[1] <- "ProbeNames"

# annotation needed for using functions
ann450k <- readRDS(file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/ann450k_GEO15sets.rds")
annEPIC <- readRDS(file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/annEPIC_GEO15sets.rds")

```


* In GEO_phenotypes_Filtered, we filter out failed samples for GEO_RGset_450k_normalPlacenta[,5] (GSM1617002_7668610068_R01C01) and GEO_RGset_450k_MaternalBlood[,8] (GSM1616993_7668610068_R06C02). These 2 failed samples are placenta sample and maternal blood sample respectively, and both of them were from first trimester.


# 2. Sample and probe filtering & Background and dye bias correction & nomalisation

There are 6 tissue types (we choose tissue types located near placenta) in total from GEO.

## A) Function for filtering

FilterFUN & use FilterFUN

TissueType: Placenta, Amnion, Chorion, Decidua, Maternal whole blood, Umbilical cord blood.

### Function `FilterFUN`


```{r FilterFUN}

source(file = "~/DNAme/Process_decidual/GEO_TechPaper/Functions/GEOdataAnalysis_FilterFUN.R")

```

** I will do nomalisation in next Rmd document, because there may be batch effects between GEO studies, I need to use BMIQ and modified BMIQ to normalise all data sets.**



## B) Use the function FilterFUN to get **filtered Methyl sets** for all the samples from 6 kinds of tissues located around placenta during early development.


### read in EPIC and 450k seperately

```{r}

GEO_baseDir <- file.path("/home/qianhui/DNAme/Process_decidual/GEO_dataSets/IDAT_all")

# get the RGset for all 392 450k samples
GEO_RGset_450k <- read.metharray.exp(base = GEO_baseDir, 
                                     targets = GEO_phenotypes_Filtered_sub[GEO_phenotypes_Filtered_sub$ArrayType%in%"450K",], 
                                     extended = TRUE, force = TRUE)

# get the RGset for all 16 EPIC samples
GEO_RGset_EPIC <- read.metharray.exp(base = GEO_baseDir, 
                                     targets = GEO_phenotypes_Filtered_sub[GEO_phenotypes_Filtered_sub$ArrayType%in%"EPIC",], 
                                     extended = TRUE, force = TRUE)

```


### get filtered Msets
```{r}

GEO_Mset_450k_Fil <- FilterFun(RGsetChara = GEO_RGset_450k,
          crossreactiveProbes = crossreactiveProbes450k, 
          annotation = ann450k, DropXreactive = TRUE, 
          DropSNP = TRUE, DropXY = TRUE)

GEO_Mset_EPIC_Fil <- FilterFun(RGsetChara = GEO_RGset_EPIC,
          crossreactiveProbes = crossreactiveProbesEPIC, 
          annotation = annEPIC, DropXreactive = TRUE,
          DropSNP = TRUE, DropXY = TRUE)

```

### get common probes
```{r}

intersetProbes <- intersect(rownames(GEO_Mset_450k_Fil), rownames(GEO_Mset_EPIC_Fil))


GEO_Mset_450k <- preprocessRaw(GEO_RGset_450k)
GEO_Mset_EPIC <- preprocessRaw(GEO_RGset_EPIC)


RemovedProbe_450k <- GEO_Mset_450k[!(rownames(GEO_Mset_450k)%in%intersetProbes),] %>% rownames()

RemovedProbe_EPIC <- GEO_Mset_EPIC[!(rownames(GEO_Mset_EPIC)%in%intersetProbes),] %>% rownames()

```


## C) Background correction and dye bias correction:

### Preprocess Enmix
```{r}

GEO_Mset_450k_Fil_enmix <- preprocessENmix(GEO_RGset_450k, exCpG = RemovedProbe_450k, nCores = 10)

GEO_Mset_EPIC_Fil_enmix <- preprocessENmix(GEO_RGset_EPIC, exCpG = RemovedProbe_EPIC, nCores = 10)

```

<!-- Preprocess quantile 1 -->
<!-- ```{r} -->

<!-- GEO_Mset_450k_Tissue_Fil_ENQN1 <- norm.quantile(GEO_Mset_450k_Tissue_Fil_enmix, method = "quantile1") -->

<!-- GEO_Mset_EPIC_Tissue_Fil_ENQN1 <- norm.quantile(GEO_Mset_EPIC_Tissue_Fil_enmix, method = "quantile1") -->

<!-- ``` -->


### Get Beta and M values from Msets
```{r}

betaList <- list()
Mlist <- List()

for (i in c("GEO_Mset_450k_Fil_enmix","GEO_Mset_EPIC_Fil_enmix")){
    Mset <- get(i)
    beta <- minfi::getBeta(Mset) 
    betaList[[i]] <- beta %>% as.data.frame() %>% rownames_to_column(var="Probe") 
    
    ## calculate M values
    M <- log2(beta/(1 - beta))
    M_v1 <- M[apply(M, 1, function(x) all(is.finite(x))), ]
    Mlist[[i]] <- M_v1 %>% as.data.frame() %>% rownames_to_column(var="Probe") 
}
   
```

### bind beta and M values from 450k and EPIC array
```{r}

Beta_all <- betaList$GEO_Mset_450k_Fil_enmix %>% 
  left_join(betaList$GEO_Mset_EPIC_Fil_enmix, by="Probe") %>% 
  column_to_rownames(var="Probe") %>% 
  as.matrix()

M_all <- Mlist$GEO_Mset_450k_Fil_enmix %>% 
  left_join(Mlist$GEO_Mset_EPIC_Fil_enmix, by="Probe") %>% 
  column_to_rownames(var="Probe") %>% 
  as.matrix()

```

PCA for data before normalisation
```{r}
PCA_Mall <- PCA(t(M_all))
```

matched phenodata for 408 samples
```{r}
Matched_pd <- GEO_phenotypes_Filtered_sub[match(colnames(M_all), GEO_phenotypes_Filtered_sub$Sample_name),]
```

clustering for data before normalisation
```{r}

library(mclust)
source(file = "~/DNAme/Process_decidual/GEO_TechPaper/Functions/GEOdataAnalysis_ggplotPCAFUN.R")

Mall_pcaRES <- ggplotPCA(PCA_Placenta_M = PCA_Mall, 
                         Matched_pd = Matched_pd, 
                         value = "Mval_allNorm")

Mall_pcaRES %>% 
  ggplot(aes(x, y)) + 
  geom_point(aes(shape = Tissue, col=Study), alpha = 1, size = 5)+ 
  # scale_color_manual(values = namedPal2)+
  scale_shape_manual(values=1:11)+
  labs(x="PC1", y="PC2")+
  theme(text = element_text(size=20))


df <- PCA_Mall$ind$coord %>% as.data.frame() %>% rownames_to_column(var="Sample_name") %>% 
  left_join(GEO_phenotypes_Filtered, by="Sample_name") %>% column_to_rownames(var="Sample_name")

a1 <- mclust::Mclust(df[,c("Dim.1","Dim.2")],2) #for 2 groups
table(a1$classification,df$Sample)
with(df,plot(Dim.2 ~ Dim.1,pch=as.numeric(as.factor(Sample)),col=a1$classification))
with(df,legend("topright",levels(as.factor(Sample)),pch=1:6))

```


## D) Normalisation:

```{r}

library(ChAMP)
BetaAllNorm <- champ.norm(beta=Beta_all,
             resultsDir="~/DNAme/Process_decidual/RDSfiles/",
             method="BMIQ",
             plotBMIQ=FALSE,
             arraytype="450K",
             cores=3)
  
  # M values
  Mnorm <- log2(BetaAllNorm/(1 - BetaAllNorm))
  MAllNorm <- Mnorm[apply(Mnorm, 1, function(x) all(is.finite(x))), , drop=FALSE]
  
  
  NormBMlist <- list(NormBAll=BetaAllNorm, NormMAll=MAllNorm, 
                     phenoDataAll=Matched_pd)
  
  saveRDS(NormBMlist, file = "/home/qianhui/DNAme/Process_decidual/RDSfiles/NormBMlist_300k_sub.rds")
  

```


PCA for data after normalisation
```{r}

PCA_all <- PCA(t(MAllNorm)) 

saveRDS(PCA_all, file  = "/home/qianhui/DNAme/Process_decidual/RDSfiles/PCA_all_sub.Rds")

```


clustering
```{r}

source(file = "~/DNAme/Process_decidual/GEO_TechPaper/Functions/GEOdataAnalysis_ggplotPCAFUN.R")

Mall_pcaRES <- ggplotPCA(PCA_Placenta_M = PCA_all, 
                         Matched_pd = Matched_pd, 
                         value = "Mval_allNorm")

Mall_pcaRES %>% 
  ggplot(aes(x, y)) + 
  geom_point(aes(shape = Tissue, col=Trimester), alpha = 1, size = 5)+ 
  scale_shape_manual(values=1:11)+
  labs(x="PC1", y="PC2")+
  theme(text = element_text(size=8))



df <- PCA_all$ind$coord %>% as.data.frame() %>% rownames_to_column(var="Sample_name") %>% 
  left_join(Matched_pd, by="Sample_name") %>% column_to_rownames(var="Sample_name")

a1 <- mclust::Mclust(df[,c("Dim.1", "Dim.2")],2) #for 2 groups
table(a1$classification,df$Sample)
with(df,plot(Dim.2 ~ Dim.1,pch=as.numeric(as.factor(Sample)),col=a1$classification))
with(df,legend("topright",levels(as.factor(Sample)),pch=1:6))

a1$z %>% as.data.frame() %>% rownames_to_column() %>% arrange(V2)


df$class <- a1$classification

PCAclassPlot <- df %>%
  ggplot(aes(Dim.1, Dim.2)) + 
  geom_point(aes(shape = Sample, col=as.factor(class)), alpha = 1, size = 2)+ 
  scale_shape_manual(values=c(0:5))+
  scale_color_manual(name="Clusters", values = c("1"="#000000","2"="#D95F02"), 
                     labels=c("1"="Cluster 1", "2"="Cluster 2"))+
  labs(x="PC1 (34.12%)", y="PC2 (12.03%)")+
  scale_y_continuous(breaks=c(-600,-300,0, 300, 600), limits = c(-600, 800))+
  scale_x_continuous(breaks=c(-400,0, 400, 800), limits = c(-410, 1150))+
  guides(color=guide_legend(nrow = 3), shape=guide_legend(nrow = 2)) +
  theme(text = element_text(size=9),
        legend.title=element_text(size=9),
        legend.text=element_text(size=9),
        axis.text.x =element_text(size=9, colour = "black"),
        axis.text.y =element_text(size=9, colour = "black"),
        legend.position="top")

```





